## Custom Change-Point Detector

## Objective

The goal of this example is to show how to integrate a custom change-point detector based on joinpoint regression.

This notebook is not only about the integration contract. It is also meant to motivate a very interpretable idea of change-point detection: sometimes the important event in a series is the point where one trend stops and another begins.

## Why this method matters

Many change-point methods are framed as optimization or segmentation problems. That is useful, but it can feel abstract for readers who think first in terms of regression lines and slope changes. Joinpoint regression gives a more visual intuition: fit piecewise linear behavior and estimate where the line changes direction or slope.

This is a strong custom example because:

- it connects change-point detection with a familiar regression interpretation;
- it shows how external model objects can be wrapped inside Harbinger;
- it motivates structural breaks as changes in trend, not only as generic break indices.

## Method at a glance

The detector first fits a linear model and then refines it with `segmented::segmented()`, which estimates one or more joinpoints. The estimated breakpoint positions are converted into Harbinger's logical event vector and then reused in the regular plotting and evaluation workflow.

Joinpoint or segmented regression is a good custom extension example because it combines a standard regression fit with an explicit estimate of the structural break location. Here we use the `segmented` package and wrap it inside the usual Harbinger interface.



## Walkthrough



### Prepare the Example

We begin by organizing the environment, loading the packages, and selecting the dataset used in the notebook. This part is intentionally more direct: the goal is to make the starting point explicit before the method-specific reasoning begins.

```{r}
# installation
# install.packages(c("harbinger", "daltoolbox", "segmented"))

library(daltoolbox)
library(harbinger)
```





### Define the Support Structures

Before applying the workflow itself, we define the helper functions or custom objects that make the example possible. This is one of the most important didactic moments in extension-oriented notebooks because it shows the contract that Harbinger expects and where the reader can adapt the behavior later.

```{r}
hcp_joinpoint_custom <- function(npsi = 1) {
  obj <- harbinger()
  obj$npsi <- npsi
  class(obj) <- append("hcp_joinpoint_custom", class(obj))
  obj
}

fit.hcp_joinpoint_custom <- function(obj, data, ...) {
  x <- seq_along(data)
  base_model <- stats::lm(data ~ x)
  obj$model <- segmented::segmented(base_model, seg.Z = ~ x, npsi = obj$npsi)
  obj
}

detect.hcp_joinpoint_custom <- function(obj, serie, ...) {
  obj <- obj$har_store_refs(obj, serie)

  if (is.null(obj$model)) {
    obj <- fit(obj, obj$serie)
  }

  cp <- rep(FALSE, length(obj$serie))
  psi <- round(obj$model$psi[, "Est."])
  psi <- psi[psi >= 1 & psi <= length(cp)]
  cp[psi] <- TRUE

  obj$har_restore_refs(obj, change_points = cp)
}
```

We now fit and use the custom joinpoint detector on a simple change-point series.





### Configure the Method

The next step is to instantiate the method and, when necessary, fit it to the selected series. This is where the notebook makes its analytical choice explicit: the parameters chosen here determine what kind of pattern the detector or transformer will become sensitive to and how the later outputs should be interpreted.

```{r}
data(examples_changepoints)
dataset <- examples_changepoints$simple

model <- hcp_joinpoint_custom(npsi = 1)
model <- fit(model, dataset$serie)
detection <- detect(model, dataset$serie)
```





### Evaluate What Was Found

After producing detections or transformed outputs, we compare them with the reference labels whenever they are available. This stage matters because it connects the visual intuition of the method with an explicit measurement of quality, helping the learner understand not only whether the method runs, but how well it behaves.

```{r}
evaluation <- evaluate(model, detection$event, dataset$event)
evaluation$confMatrix
```





### Interpret the Result Visually

The final plots are not just illustrations. They help the reader connect the method's internal output with the original series, making it easier to see why a point, range, motif, or symbolic pattern was emphasized and whether that emphasis is coherent with the stated objective of the example.

```{r}
har_plot(model, dataset$serie, detection, dataset$event)
```

This example highlights the usual role of a custom change-point extension in Harbinger: fit a structural model, convert the break estimate to the common detection table, and then reuse the regular evaluation and plotting workflow.

## References

- Muggeo, V. M. R. (2003). Estimating regression models with unknown break-points. Statistics in Medicine, 22(19), 3055-3071.
