## Objective

FBIAD (Forward and Backward Inertial Anomaly Detector) compares deviations from sliding-window means computed forward and backward in time, then merges evidence. In this tutorial we:

- Load and visualize a simple anomaly dataset
- Configure and run the FBIAD detector (`hanr_fbiad`)
- Inspect detections, evaluate, and plot residual magnitudes

## Method at a glance

FBIAD regression anomaly detection: Forward and Backward Inertial Anomaly Detector compares each point against forward and backward inertia, flagging observations that break both temporal tendencies. Scores are summarized and thresholded using `harutils()`.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure the FBIAD detector
model <- hanr_fbiad()
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

## References

- Lima, J., et al. Forward and Backward Inertial Anomaly Detector: A Novel Time Series Event Detection Method. IJCNN, 2022. doi:10.1109/IJCNN55064.2022.9892088
