## Objective

The GARCH-based detector models conditional heteroskedasticity and flags large standardized residuals as anomalies. We will:

- Load and visualize a sample anomaly dataset
- Configure and run the GARCH detector (`hanr_garch`)
- Inspect detections, evaluate, and plot residuals with thresholds

## Method at a glance

GARCH-based regression anomaly detection: This detector estimates a GARCH model to capture conditional heteroskedasticity and flags observations with large standardized residuals as anomalies. A common configuration is sGARCH(1,1) with an ARMA mean. Residuals are summarized and thresholded using `harutils()`.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger








### Prepare the Example

This setup anchors the notebook in the specific series used to examine `hanr_garch`. The semantic point is the one stated above: gARCH-based regression anomaly detection: This detector estimates a GARCH model to capture conditional heteroskedasticity and flags observations with large standardized residuals as anomalies, so the raw signal needs to be visible before any fitting step hides that structure behind model output.

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```





```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```





```{r}
# Load example anomaly datasets
data(examples_anomalies)
```



```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```







### Interpret the Result Visually

This first visual pass establishes what the method should react to in the raw series. Keep the method summary in mind here, because gARCH-based regression anomaly detection: This detector estimates a GARCH model to capture conditional heteroskedasticity and flags observations with large standardized residuals as anomalies and the plot tells you whether that structure is clean, weak, local, repeated, or mixed with other effects.

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```







### Configure the Method

The choices below turn the central modeling idea into concrete parameters. They matter because gARCH-based regression anomaly detection: This detector estimates a GARCH model to capture conditional heteroskedasticity and flags observations with large standardized residuals as anomalies, so each argument controls how strongly the method will emphasize that pattern when it later produces anomaly flags.

```{r}
# Configure the GARCH-based detector
model <- hanr_garch()
```



```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```







### Run the Core Analysis

This is the moment where the notebook tests its central assumption on actual data. After applying `hanr_garch`, the important question is whether the resulting anomaly flags really correspond to the pattern implied by the method description above, rather than to arbitrary numerical variation.

```{r}
# Run detection to compute residual magnitudes and flags
detection <- detect(model, dataset$serie)
```



```{r}
# Show detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```







### Evaluate What Was Found

The evaluation asks whether the anomaly flags produced by `hanr_garch` match the labeled structure on this dataset. Read the scores as evidence about the method's assumptions in practice, not as detached summary numbers.

```{r}
# Evaluate detections against labeled events
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```







### Interpret the Result Visually

This visual check puts the model output back on top of the original signal. What matters now is whether the highlighted anomaly flags line up with the structure suggested by the method, which is the real semantic test of whether the example is teaching the right lesson.

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```



```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

## References

- Engle, R. F. (1982). Autoregressive Conditional Heteroscedasticity with Estimates of the Variance of United Kingdom Inflation. Econometrica, 50(4), 987-1007.
- Bollerslev, T. (1986). Generalized Autoregressive Conditional Heteroskedasticity. Journal of Econometrics, 31(3), 307-327.