## Objective

This Rmd shows a simple supervised baseline for anomaly classification using `hanc_ml` with a Majority Classifier (`cla_majority`). It predicts the most frequent class in training. Steps: load packages/data, visualize, preprocess (split + normalize), define and fit, detect, evaluate, and plot results.

## Method at a glance

Majority-class classification anomaly detector: Supervised anomaly detection using a majority-class baseline to illustrate the `hanc_ml` interface. Events correspond to predicted positive class above the threshold.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger








### Prepare the Example

This setup anchors the notebook in the specific series used to examine `cla_majority`. The semantic point is the one stated above: majority-class classification anomaly detector: Supervised anomaly detection using a majority-class baseline to illustrate the `hanc_ml` interface, so the raw signal needs to be visible before any fitting step hides that structure behind model output.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```





```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```





```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```



```{r}
# Use the "tt" time series (labeled)
dataset <- examples_anomalies$tt

head(dataset)
```







### Interpret the Result Visually

This first visual pass establishes what the method should react to in the raw series. Keep the method summary in mind here, because majority-class classification anomaly detector: Supervised anomaly detection using a majority-class baseline to illustrate the `hanc_ml` interface and the plot tells you whether that structure is clean, weak, local, repeated, or mixed with other effects.

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```







### Configure the Method

The choices below turn the central modeling idea into concrete parameters. They matter because majority-class classification anomaly detector: Supervised anomaly detection using a majority-class baseline to illustrate the `hanc_ml` interface, so each argument controls how strongly the method will emphasize that pattern when it later produces event classifications.

```{r}
# Data preprocessing: split and scale


train <- dataset[1:80,]
test <- dataset[-(1:80),]

norm <- minmax()
norm <- fit(norm, train)
train_n <- transform(norm, train)
summary(train_n)
```



```{r}
# Define Majority Classifier for events (hanc_ml + cla_majority)
model <- hanc_ml(cla_majority("event", c("FALSE", "TRUE")))
```



```{r}
# Fit the model on training data
model <- fit(model, train_n)
detection <- detect(model, train_n)
print(detection |> dplyr::filter(event==TRUE))
# Evaluate training performance
evaluation <- evaluate(model, detection$event, as.logical(train_n$event))
print(evaluation$confMatrix)
```







### Interpret the Result Visually

This visual check puts the model output back on top of the original signal. What matters now is whether the highlighted event classifications line up with the structure suggested by the method, which is the real semantic test of whether the example is teaching the right lesson.

```{r}
# Plot training detections
  har_plot(model, train_n$serie, detection, as.logical(train_n$event))
```







### Run the Core Analysis

This is the moment where the notebook tests its central assumption on actual data. After applying `cla_majority`, the important question is whether the resulting event classifications really correspond to the pattern implied by the method description above, rather than to arbitrary numerical variation.

```{r}
# Prepare test data (apply same scaler)
  test_n <- transform(norm, test)
```



```{r}
# Detect and evaluate on test data
  detection <- detect(model, test_n)

  print(detection |> dplyr::filter(event==TRUE))

  evaluation <- evaluate(model, detection$event, as.logical(test_n$event))
  print(evaluation$confMatrix)
```







### Interpret the Result Visually

This visual check puts the model output back on top of the original signal. What matters now is whether the highlighted event classifications line up with the structure suggested by the method, which is the real semantic test of whether the example is teaching the right lesson.

```{r}
# Plot test detections
  har_plot(model, test_n$serie, detection, as.logical(test_n$event))
```



```{r}
# Plot residual scores and threshold
har_plot(model, attr(detection, "res"), detection, test_n$event, yline = attr(detection, "threshold"))
```

## References

- Bishop, C. M. (2006). Pattern Recognition and Machine Learning. Springer.
- Hyndman, R. J., Athanasopoulos, G. (2021). Forecasting: Principles and Practice. OTexts.