## Objective

This tutorial shows supervised anomaly detection with an SVM classifier over a labeled train/test split. 

Steps:
- Load and visualize the dataset
- Normalize, train SVM, evaluate on train and test
- Plot detections and residual magnitudes

## Method at a glance

SVM classification anomaly detector: Supervised anomaly detection with an SVM classifier trained on labeled events; positive-class probabilities above a threshold correspond to detected events.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough







### Prepare the Example

We begin by organizing the environment, loading the packages, and selecting the dataset used in the notebook. This part is intentionally more direct: the goal is to make the starting point explicit before the method-specific reasoning begins.

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```





```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```





```{r}
# Load example anomaly datasets
data(examples_anomalies)
```



```{r}
# Select the train/test dataset
dataset <- examples_anomalies$tt

head(dataset)
```







### Interpret the Result Visually

The final plots are not just illustrations. They help the reader connect the method's internal output with the original series, making it easier to see why a point, range, motif, or symbolic pattern was emphasized and whether that emphasis is coherent with the stated objective of the example.

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```







### Configure the Method

The next step is to instantiate the method and, when necessary, fit it to the selected series. This is where the notebook makes its analytical choice explicit: the parameters chosen here determine what kind of pattern the detector or transformer will become sensitive to and how the later outputs should be interpreted.

```{r}
# Split into train/test and normalize features
train <- dataset[1:80,]
test <- dataset[-(1:80),]

norm <- minmax()
norm <- fit(norm, train)
train_n <- transform(norm, train)
summary(train_n)
```



```{r}
# Configure SVM classifier
model <- hanc_ml(cla_svm("event", c("FALSE", "TRUE"), epsilon = 0.0, cost = 20.000))
```



```{r}
# Fit on training data and evaluate on train
model <- fit(model, train_n)
detection <- detect(model, train_n)
print(detection |> dplyr::filter(event == TRUE))
evaluation <- evaluate(model, detection$event, as.logical(train_n$event))
print(evaluation$confMatrix)
```







### Interpret the Result Visually

The final plots are not just illustrations. They help the reader connect the method's internal output with the original series, making it easier to see why a point, range, motif, or symbolic pattern was emphasized and whether that emphasis is coherent with the stated objective of the example.

```{r}
# Plot training detections
har_plot(model, train_n$serie, detection, as.logical(train_n$event))
  
```







### Run the Core Analysis

With the environment and the method ready, we execute the central analytical step and inspect its immediate output. This is the point where the abstract idea described earlier becomes operational, so the reader should pay attention to what is produced and how Harbinger standardizes the result.

```{r}
# Prepare normalized test set
test_n <- transform(norm, test)
```



```{r}
# Detect and evaluate on test
detection <- detect(model, test_n)
print(detection |> dplyr::filter(event == TRUE))
evaluation <- evaluate(model, detection$event, as.logical(test_n$event))
print(evaluation$confMatrix)
```







### Interpret the Result Visually

The final plots are not just illustrations. They help the reader connect the method's internal output with the original series, making it easier to see why a point, range, motif, or symbolic pattern was emphasized and whether that emphasis is coherent with the stated objective of the example.

```{r}
# Plot test detections
har_plot(model, test_n$serie, detection, as.logical(test_n$event))
  
```



```{r}
# Plot residual magnitude and decision threshold
har_plot(model, attr(detection, "res"), detection, test_n$event, yline = attr(detection, "threshold"))
  
```

## References

- Bishop, C. M. (2006). Pattern Recognition and Machine Learning. Springer.
