## Objective

This notebook demonstrates anomaly detection with a convolutional autoencoder (`han_autoencoder(..., autoenc_conv_ed, ...)`). The model reconstructs windows via Conv1D layers; high reconstruction error indicates anomalies. Steps: load packages/data, visualize, define the architecture/epochs, fit, detect, evaluate, and plot.

## Method at a glance

Convolutional autoencoder (encode-decode): A convolutional autoencoder reconstructs windows via Conv1D layers; high reconstruction error indicates anomalies. Convolutions capture local patterns and translation invariance; thresholds use `harutils()`.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(daltoolboxdp)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```

```{r}
# Select a simple synthetic time series with labeled anomalies
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define convolutional autoencoder-based detector (autoenc_conv_ed)
  model <- han_autoencoder(3, 2, autoenc_conv_ed, num_epochs = 1500)
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect anomalies (reconstruction error -> events)
  detection <- detect(model, dataset$serie)
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
```
```{r}
# Plot residual scores and threshold
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

## References

- Sakurada, M., Yairi, T. (2014). Anomaly Detection Using Autoencoders with Nonlinear Dimensionality Reduction. MLSDA 2014.
