## Objective

This notebook demonstrates anomaly detection with an LSTM autoencoder (`han_autoencoder(..., autoenc_lstm_ed, ...)`). The model encodes and decodes sequences; high reconstruction error flags anomalies. Steps: load packages/data, visualize, define the architecture/epochs, fit, detect, evaluate, and plot.

## Method at a glance

LSTM autoencoder (encode-decode): An LSTM autoencoder learns temporal dependencies by encoding and decoding sequences; large reconstruction error flags anomalies. The model optimizes sequence reconstruction and thresholds errors via `harutils()`.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger








### Prepare the Example

This setup anchors the notebook in the specific series used to examine `han_autoencoder(..., autoenc_lstm_ed, ...)`. The semantic point is the one stated above: lSTM autoencoder (encode-decode): An LSTM autoencoder learns temporal dependencies by encoding and decoding sequences; large reconstruction error flags anomalies, so the raw signal needs to be visible before any fitting step hides that structure behind model output.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```





```{r}
# Load required packages
library(daltoolbox)
library(daltoolboxdp)
library(harbinger) 
```





```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```



```{r}
# Select a simple synthetic time series with labeled anomalies
dataset <- examples_anomalies$simple
head(dataset)
```







### Interpret the Result Visually

This first visual pass establishes what the method should react to in the raw series. Keep the method summary in mind here, because lSTM autoencoder (encode-decode): An LSTM autoencoder learns temporal dependencies by encoding and decoding sequences; large reconstruction error flags anomalies and the plot tells you whether that structure is clean, weak, local, repeated, or mixed with other effects.

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```







### Configure the Method

The choices below turn the central modeling idea into concrete parameters. They matter because lSTM autoencoder (encode-decode): An LSTM autoencoder learns temporal dependencies by encoding and decoding sequences; large reconstruction error flags anomalies, so each argument controls how strongly the method will emphasize that pattern when it later produces reconstruction-based anomaly flags.

```{r}
# Define LSTM autoencoder-based detector (autoenc_lstm_ed)
  model <- han_autoencoder(3, 2, autoenc_lstm_ed, num_epochs = 1500)
```



```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```







### Run the Core Analysis

This is the moment where the notebook tests its central assumption on actual data. After applying `han_autoencoder(..., autoenc_lstm_ed, ...)`, the important question is whether the resulting reconstruction-based anomaly flags really correspond to the pattern implied by the method description above, rather than to arbitrary numerical variation.

```{r}
# Detect anomalies (reconstruction error -> events)
  detection <- detect(model, dataset$serie)
```



```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```







### Evaluate What Was Found

The evaluation asks whether the reconstruction-based anomaly flags produced by `han_autoencoder(..., autoenc_lstm_ed, ...)` match the labeled structure on this dataset. Read the scores as evidence about the method's assumptions in practice, not as detached summary numbers.

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```







### Interpret the Result Visually

This visual check puts the model output back on top of the original signal. What matters now is whether the highlighted reconstruction-based anomaly flags line up with the structure suggested by the method, which is the real semantic test of whether the example is teaching the right lesson.

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
```



```{r}
# plotting the residuals
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

## References

- Sakurada, M., Yairi, T. (2014). Anomaly Detection Using Autoencoders with Nonlinear Dimensionality Reduction. MLSDA 2014.