LSTM autoencoder (encode-decode): An LSTM autoencoder learns temporal dependencies by encoding and decoding sequences; large reconstruction error flags anomalies. The model optimizes sequence reconstruction and thresholds errors via `harutils()`.

Objectives: 
This Rmd demonstrates anomaly detection with an LSTM autoencoder (`han_autoencoder(..., autoenc_lstm_ed, ...)`). The model encodes and decodes sequences; high reconstruction error flags anomalies. Steps: load packages/data, visualize, define the architecture/epochs, fit, detect, evaluate, and plot.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(daltoolboxdp)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```

```{r}
# Select a simple synthetic time series with labeled anomalies
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define LSTM autoencoder-based detector (autoenc_lstm_ed)
  model <- han_autoencoder(3, 2, autoenc_lstm_ed, num_epochs = 1500)
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect anomalies (reconstruction error -> events)
  detection <- detect(model, dataset$serie)
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
```
```{r}
# plotting the residuals
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```


References 
- Sakurada, M., Yairi, T. (2014). Anomaly Detection Using Autoencoders with Nonlinear Dimensionality Reduction. MLSDA 2014.
