Autoencoder (encode-decode): An autoencoder reconstructs sliding windows; large reconstruction errors indicate anomalies. This example uses a feed-forward encoder-decoder. Training minimizes reconstruction loss; detection thresholds use `harutils()`.

This Rmd demonstrates anomaly detection via a basic feed-forward autoencoder (`han_autoencoder(..., autoenc_ed, ...)`). The model reconstructs the input window; high reconstruction error indicates anomalies. Steps: load packages/data, visualize, define the autoencoder (layers/epochs), fit, detect, evaluate, and plot series and residuals.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(daltoolboxdp)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```

```{r}
# Select a simple synthetic time series with labeled anomalies
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define autoencoder-based detector (autoenc_ed)
# - first/second args: encoder/decoder sizes; num_epochs: training epochs
model <- han_autoencoder(3, 2, autoenc_ed, num_epochs = 1500)
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect anomalies (reconstruction error -> events)
  detection <- detect(model, dataset$serie)
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual scores and threshold
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 

- Sakurada, M., Yairi, T. (2014). Anomaly Detection Using Autoencoders with Nonlinear Dimensionality Reduction. MLSDA 2014.
