Naive Bayes classification anomaly detector: Supervised anomaly detection using Naive Bayes classification on labeled sequences; positive-class probabilities above the threshold indicate events.

This tutorial shows supervised anomaly detection via Naive Bayes classification over a labeled train/test split. We will:

- Load a train/test dataset and visualize it
- Normalize features, train a Naive Bayes classifier, and evaluate on train and test
- Plot detections and residual magnitudes

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select the train/test dataset
dataset <- examples_anomalies$tt

head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Split into train/test and normalize features
train <- dataset[1:80,]
test <- dataset[-(1:80),]

norm <- minmax()
norm <- fit(norm, train)
train_n <- transform(norm, train)
summary(train_n)
```

```{r}
# Configure NB classifier
model <- hanc_ml(cla_nb("event", c("FALSE", "TRUE")))
```

```{r}
# Fit on training data, evaluate on train
model <- fit(model, train_n)
detection <- detect(model, train_n)
print(detection |> dplyr::filter(event == TRUE))
evaluation <- evaluate(model, detection$event, as.logical(train_n$event))
print(evaluation$confMatrix)
```

```{r}
# Plot training detections
har_plot(model, train_n$serie, detection, as.logical(train_n$event))
```

```{r}
# Prepare normalized test set
test_n <- transform(norm, test)
```

```{r}
# Detect and evaluate on test
detection <- detect(model, test_n)
print(detection |> dplyr::filter(event == TRUE))
evaluation <- evaluate(model, detection$event, as.logical(test_n$event))
print(evaluation$confMatrix)
```

```{r}
# Plot test detections
har_plot(model, test_n$serie, detection, as.logical(test_n$event))
```

```{r}
# Plot residual magnitude and decision threshold
har_plot(model, attr(detection, "res"), detection, test_n$event, yline = attr(detection, "threshold"))
```


References 
- Bishop, C. M. (2006). Pattern Recognition and Machine Learning. Springer.
