FFT AMOC CUSUM regression anomaly detection: This hybrid detector computes the FFT and transforms the power spectrum with CUSUM to emphasize gradual spectral shifts. AMOC selects a single changepoint that serves as a frequency cutoff. Components below the cutoff are removed, the inverse FFT reconstructs a high-pass signal, and residual magnitudes are thresholded via `harutils()`.

Objectives: This Rmd demonstrates a hybrid residual-based anomaly detector combining FFT preprocessing, AMOC (At Most One Change) modeling, and CUSUM scoring via `hanr_fft_amoc_cusum()`. The pipeline models dominant frequencies, captures a single change, and accumulates deviations to flag anomalies. Steps: load packages/data, visualize, define and fit, detect, evaluate, and plot.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```

```{r}
# Select a simple synthetic time series with labeled anomalies
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define FFT + AMOC + CUSUM detector
  model <- hanr_fft_amoc_cusum()
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect anomalies
  detection <- detect(model, dataset$serie)
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
  
```

```{r}
# Plot residual scores and threshold
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
  
```

References 
- Sobrinho, E. P., et al. Fine-Tuning Detection Criteria for Enhancing Anomaly Detection in Time Series. SBBD, 2025. doi:10.5753/sbbd.2025.247063
