RTAD regression anomaly detector: RTAD adapts to local dynamics using EMD-derived components and robust dispersion within sliding windows. Points with large deviations relative to the modeled behavior are flagged as anomalies; thresholds use `harutils()`.

Objectives: 
This Rmd applies the Residual Error-based Detector (`hanr_rt`) to identify anomalies in a univariate time series. The pipeline fits a baseline model, scores residuals, and flags points exceeding an adaptive threshold. Steps: load packages/data, visualize, define and fit the model, detect, evaluate, and plot results and residuals.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```

```{r}
# Select a simple synthetic time series with labeled anomalies
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define Residual Error-based Detector (hanr_rt)
  model <- hanr_rtad()
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect anomalies (compute residuals and events)
  detection <- detect(model, dataset$serie)
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual scores and threshold
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3

