FBIAD regression anomaly detection: Forward and Backward Inertial Anomaly Detector compares each point against forward and backward inertia, flagging observations that break both temporal tendencies. Scores are summarized and thresholded using `harutils()`.

FBIAD (Forward and Backward Inertial Anomaly Detector) compares deviations from sliding-window means computed forward and backward in time, then merges evidence. In this tutorial we:

- Load and visualize a simple anomaly dataset
- Configure and run the FBIAD detector (`hanr_fbiad`)
- Inspect detections, evaluate, and plot residual magnitudes

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure the FBIAD detector
model <- hanr_fbiad()
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 
- Lima, J., et al. Forward and Backward Inertial Anomaly Detector: A Novel Time Series Event Detection Method. IJCNN, 2022. doi:10.1109/IJCNN55064.2022.9892088
