Wavelet regression anomaly detection: Multi-resolution analysis via MODWT wavelet decomposition; detail coefficients are aggregated to form a magnitude signal. Large localized energy indicates anomalies, with thresholding via `harutils()`.

The wavelet-based detector decomposes the series into multi-resolution detail bands (MODWT) and flags large aggregated high-frequency energy as anomalies. Steps:

- Load and visualize a simple anomaly dataset
- Configure and run `hanr_wavelet`
- Inspect detections, evaluate, and plot residual magnitudes and thresholds

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure the wavelet-based detector
model <- hanr_wavelet()
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 
- Mallat, S. (1989). A theory for multiresolution signal decomposition: The wavelet representation. IEEE TPAMI, 11(7), 674–693.

