GARCH-based regression anomaly detection: This detector estimates a GARCH model to capture conditional heteroskedasticity and flags observations with large standardized residuals as anomalies. A common configuration is sGARCH(1,1) with an ARMA mean. Residuals are summarized and thresholded using `harutils()`.

The GARCH-based detector models conditional heteroskedasticity and flags large standardized residuals as anomalies. We will:

- Load and visualize a sample anomaly dataset
- Configure and run the GARCH detector (`hanr_garch`)
- Inspect detections, evaluate, and plot residuals with thresholds

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure the GARCH-based detector
model <- hanr_garch()
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection to compute residual magnitudes and flags
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labeled events
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 
- Engle, R. F. (1982). Autoregressive Conditional Heteroscedasticity with Estimates of the Variance of United Kingdom Inflation. Econometrica, 50(4), 987–1007.
- Bollerslev, T. (1986). Generalized Autoregressive Conditional Heteroskedasticity. Journal of Econometrics, 31(3), 307–327.
