Ensemble anomaly detection: Majority-vote ensemble across multiple detectors, optionally with temporal fuzzification to combine nearby detections into a single event. Each base detector contributes a vote; final events reflect consensus.

Objectives: 
This Rmd demonstrates anomaly detection via an ensemble of residual-based detectors using `har_ensemble(...)`. Multiple base detectors (e.g., FBIAD, ARIMA, EMD) vote to produce a final event decision. Steps: load packages/data, visualize, define ensemble, fit, detect, evaluate, and plot results and residuals.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_anomalies)
```

```{r}
# Select a simple synthetic time series with labeled anomalies
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define the ensemble of base detectors
  model <- har_ensemble(hanr_fbiad(), hanr_arima(), hanr_emd())
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect anomalies via ensemble voting
  detection <- detect(model, dataset$serie)
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
  
```
```{r}
# Plot residual scores and threshold
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
  
```

References 
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
