FFT Binseg CUSUM regression anomaly detection: FFT-based filtering with a cutoff determined by applying CUSUM to the spectrum and locating a changepoint via Binary Segmentation. Low-frequency energy below the cutoff is removed; anomalies are flagged from residual magnitudes and thresholded with `harutils()`.

This variant applies CUSUM to the spectrum to emphasize aggregated changes, then uses BinSeg to choose the cutoff before high-pass filtering. 

Steps:
- Load and visualize a simple anomaly dataset
- Configure and run `hanr_fft_binseg_cusum`
- Inspect detections, evaluate, and plot residual magnitudes and thresholds

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure the FFT+CUSUM+BinSeg detector
model <- hanr_fft_binseg_cusum()
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 
- Sobrinho, E. P., et al. Fine-Tuning Detection Criteria for Enhancing Anomaly Detection in Time Series. SBBD, 2025. doi:10.5753/sbbd.2025.247063

