PCA-based regression anomaly detection: Projects multivariate observations onto principal components and reconstructs data; large reconstruction errors indicate anomalies. Data are standardized, PCA is fitted, and reconstruction errors are thresholded.

The PCA-based detector flags large reconstruction errors when projecting multivariate data onto principal components. In this tutorial we:

- Load a multivariate example and create synthetic event labels
- Fit the PCA detector on two dimensions
- Visualize detections over different series and inspect residual magnitudes

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
library(ggplot2)
```

```{r}
# Load a multivariate example and define event labels (for demo)
data("examples_harbinger")
dataset <- examples_harbinger$multidimensional
dataset$event <- FALSE
dataset$event[c(101,128,167)] <- TRUE
```

```{r}
head(dataset)
```

```{r}
# Plot the target series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Plot the second dimension
har_plot(harbinger(), dataset$x)
```

```{r}
# Fit the PCA detector on the first two columns and run detection
model <- fit(hmu_pca(), dataset[,1:2])
detection <- detect(model, dataset[,1:2])
```

```{r}
# Plot detections on the target series
grf <- har_plot(model, dataset$serie, detection, dataset$event)
grf <- grf + ylab("serie")

```

```{r}
# Plot detections on the second dimension
grf <- har_plot(model, dataset$x, detection, dataset$event)
grf <- grf + ylab("x")

```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
  
```

References 
- Jolliffe, I. T. (2002). Principal Component Analysis. Springer.
