## Objective

This notebook demonstrates `trans_xsax()`, an extended symbolic transformation for time series. The goal is to show how XSAX increases symbolic resolution relative to SAX by using a larger alphabet.

## Method at a glance

XSAX follows the same general idea as SAX: normalize the series and discretize it into symbols. The difference is that XSAX uses a larger alphanumeric alphabet, which allows a finer symbolic partition of the values.

## What you will do

- load an example motif series from Harbinger
- inspect the original numeric sequence
- configure an XSAX transformer
- transform the series into a richer symbolic encoding
- compare the numeric values with the symbolic codes

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger)
```

```{r}
# Load example motif datasets
data(examples_motifs)
```

```{r}
# Select a simple motif dataset
dataset <- examples_motifs$simple
head(dataset)
```

```{r}
# Plot the original time series
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

```{r}
# Configure the XSAX transformer
model <- trans_xsax(alpha = 16)
```

```{r}
# Fit and transform the numeric series into an extended symbolic representation
model <- fit(model, dataset$serie)
xsax_series <- transform(model, dataset$serie)
```

```{r}
# Inspect the first symbolic values
head(xsax_series, 20)
```

```{r}
# Compare original values with XSAX symbols
comparison <- data.frame(
  idx = seq_along(dataset$serie),
  value = dataset$serie,
  xsax = xsax_series
)
head(comparison, 20)
```

## References

- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
