## Objective

This notebook demonstrates the `trans_sax()` transformation. The goal is to show how a numeric time series can be converted into a symbolic representation that is easier to compare, summarize, and use in motif or discord analysis.

## Method at a glance

SAX, or Symbolic Aggregate approXimation, z-normalizes a series and discretizes it into a finite alphabet. Instead of working directly with the raw values, we work with symbolic codes that preserve coarse shape information.

## What you will do

- load a motif example series from Harbinger
- inspect the original numeric signal
- configure a SAX transformer with a chosen alphabet size
- transform the series into symbols
- compare the numeric and symbolic views of the same signal

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger)
```

```{r}
# Load example motif datasets
data(examples_motifs)
```

```{r}
# Select a simple motif dataset
dataset <- examples_motifs$simple
head(dataset)
```

```{r}
# Plot the original time series
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

```{r}
# Configure the SAX transformer
model <- trans_sax(alpha = 8)
```

```{r}
# Fit and transform the numeric series into symbols
model <- fit(model, dataset$serie)
sax_series <- transform(model, dataset$serie)
```

```{r}
# Inspect the first symbolic values
head(sax_series, 20)
```

```{r}
# Compare original values with SAX symbols
comparison <- data.frame(
  idx = seq_along(dataset$serie),
  value = dataset$serie,
  sax = sax_series
)
head(comparison, 20)
```

## References

- Lin, J., Keogh, E., Lonardi, S., Chiu, B. (2007). A symbolic representation of time series, with implications for streaming algorithms. Data Mining and Knowledge Discovery, 15, 107-144.
