## Objective

AMOC (At Most One Change) detects a single, most significant change point in a univariate time series. In this tutorial we will:

- Load a synthetic dataset with ground-truth change points
- Visualize the series
- Configure and run the AMOC detector (`hcp_amoc`)
- Inspect detections and evaluate against ground truth
- Plot the detections over the series

## Method at a glance

AMOC: AMOC targets a single most significant change point in a univariate series by optimizing a cost function over all possible change locations. It is appropriate when at most one structural break is expected. This wraps the AMOC method from the `changepoint` package.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger








### Prepare the Example

This setup anchors the notebook in the specific series used to examine `hcp_amoc`. The semantic point is the one stated above: aMOC: AMOC targets a single most significant change point in a univariate series by optimizing a cost function over all possible change locations, so the raw signal needs to be visible before any fitting step hides that structure behind model output.

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```





```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```





```{r}
# Load example change-point datasets
data(examples_changepoints)
```



```{r}
# Select a dataset ("complex" contains multiple regimes)
dataset <- examples_changepoints$complex
head(dataset)
```







### Interpret the Result Visually

This first visual pass establishes what the method should react to in the raw series. Keep the method summary in mind here, because aMOC: AMOC targets a single most significant change point in a univariate series by optimizing a cost function over all possible change locations and the plot tells you whether that structure is clean, weak, local, repeated, or mixed with other effects.

```{r}
# Plot the time series to visualize regime changes
har_plot(harbinger(), dataset$serie)
```







### Configure the Method

The choices below turn the central modeling idea into concrete parameters. They matter because aMOC: AMOC targets a single most significant change point in a univariate series by optimizing a cost function over all possible change locations, so each argument controls how strongly the method will emphasize that pattern when it later produces change-point candidates.

```{r}
# Configure the AMOC change-point detector (single change)
model <- hcp_amoc()
```



```{r}
# Fit the detector (no training required, keeps parameters on object)
model <- fit(model, dataset$serie)
```







### Run the Core Analysis

This is the moment where the notebook tests its central assumption on actual data. After applying `hcp_amoc`, the important question is whether the resulting change-point candidates really correspond to the pattern implied by the method description above, rather than to arbitrary numerical variation.

```{r}
# Run detection over the full series
detection <- detect(model, dataset$serie)
```



```{r}
# Show detected change-point indices
print(detection |> dplyr::filter(event == TRUE))
```







### Evaluate What Was Found

The evaluation asks whether the change-point candidates produced by `hcp_amoc` match the labeled structure on this dataset. Read the scores as evidence about the method's assumptions in practice, not as detached summary numbers.

```{r}
# Evaluate detections against the labeled events
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```







### Interpret the Result Visually

This visual check puts the model output back on top of the original signal. What matters now is whether the highlighted change-point candidates line up with the structure suggested by the method, which is the real semantic test of whether the example is teaching the right lesson.

```{r}
# Plot detections and ground truth on top of the series
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Hinkley, D. V. (1970). Inference about the change-point in a sequence of random variables. Biometrika, 57(1), 1-17. doi:10.1093/biomet/57.1.1
- Killick, R., Fearnhead, P., Eckley, I. A. (2012). Optimal detection of changepoints with a linear computational cost. Journal of the American Statistical Association, 107(500), 1590-1598.