## Objective

PELT (Pruned Exact Linear Time) finds multiple change points efficiently by pruning candidates under a penalized cost. In this tutorial we:

- Load and visualize a simple change-point dataset
- Configure and run the PELT detector (`hcp_pelt`)
- Inspect detections, evaluate, and plot results

## Method at a glance

PELT: PELT performs optimal partitioning of the time series under a penalized cost function while pruning candidate change locations to achieve near-linear time under suitable penalties. In Harbinger this wraps the `changepoint` package implementation and returns detected change indices along with evaluation helpers.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example change-point datasets
data(examples_changepoints)
```

```{r}
# Select the simple dataset
dataset <- examples_changepoints$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure the PELT detector
model <- hcp_pelt()
```

```{r}
# Fit the detector (no training required)
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected change points
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Killick, R., Fearnhead, P., Eckley, I. A. (2012). Optimal detection of changepoints with a linear computational cost. Journal of the American Statistical Association, 107(500), 1590–1598.
