## Objective

Generalized Fluctuation Tests (GFT) assess structural changes by examining stability of regression fits across time. In this tutorial we:

- Load and visualize a simple change-point dataset
- Configure and run the GFT-based detector (`hcp_gft`)
- Inspect detections, evaluate, and plot results

## Method at a glance

GFT: Generalized fluctuation tests assess stability of regression parameters over time using `strucchange::breakpoints()`, returning estimated break dates under information criteria.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger








### Prepare the Example

This setup anchors the notebook in the specific series used to examine `hcp_gft`. The semantic point is the one stated above: gFT: Generalized fluctuation tests assess stability of regression parameters over time using `strucchange::breakpoints()`, returning estimated break dates under information criteria, so the raw signal needs to be visible before any fitting step hides that structure behind model output.

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```





```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```





```{r}
# Load example change-point datasets
data(examples_changepoints)
```



```{r}
# Select the simple dataset
dataset <- examples_changepoints$simple
head(dataset)
```







### Interpret the Result Visually

This first visual pass establishes what the method should react to in the raw series. Keep the method summary in mind here, because gFT: Generalized fluctuation tests assess stability of regression parameters over time using `strucchange::breakpoints()`, returning estimated break dates under information criteria and the plot tells you whether that structure is clean, weak, local, repeated, or mixed with other effects.

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```







### Configure the Method

The choices below turn the central modeling idea into concrete parameters. They matter because gFT: Generalized fluctuation tests assess stability of regression parameters over time using `strucchange::breakpoints()`, returning estimated break dates under information criteria, so each argument controls how strongly the method will emphasize that pattern when it later produces change-point candidates.

```{r}
# Configure the GFT detector
model <- hcp_gft()
```



```{r}
# Fit the detector (no training required)
model <- fit(model, dataset$serie)
```







### Run the Core Analysis

This is the moment where the notebook tests its central assumption on actual data. After applying `hcp_gft`, the important question is whether the resulting change-point candidates really correspond to the pattern implied by the method description above, rather than to arbitrary numerical variation.

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```



```{r}
# Show detected change points
print(detection |> dplyr::filter(event == TRUE))
```







### Evaluate What Was Found

The evaluation asks whether the change-point candidates produced by `hcp_gft` match the labeled structure on this dataset. Read the scores as evidence about the method's assumptions in practice, not as detached summary numbers.

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```







### Interpret the Result Visually

This visual check puts the model output back on top of the original signal. What matters now is whether the highlighted change-point candidates line up with the structure suggested by the method, which is the real semantic test of whether the example is teaching the right lesson.

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Zeileis, A., Leisch, F., Kleiber, C., Hornik, K. (2002). strucchange: An R package for testing for structural change in linear regression models. Journal of Statistical Software, 7(2). doi:10.18637/jss.v007.i02
- Zeileis, A., Kleiber, C., Kramer, W., Hornik, K. (2003). Testing and dating of structural changes in practice. Computational Statistics & Data Analysis, 44(1), 109-123.