## Objective

This notebook demonstrates change-point detection using Conformal Forecasting with a Linear Regression forecaster (`hcp_cf_lr`). The method scores deviations from short-term predictions and flags structural changes when nonconformity exceeds a learned threshold. Steps: load packages/data, visualize, define model (window size), fit, detect, evaluate, and plot both detections and residuals.

## Method at a glance

ChangeFinder with linear regression: ChangeFinder with linear regression models residual deviations and applies a second-stage smoothing/thresholding to expose structural changes. Thresholding is done via `harutils()`.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger








### Prepare the Example

This setup anchors the notebook in the specific series used to examine `hcp_cf_lr`. The semantic point is the one stated above: changeFinder with linear regression: ChangeFinder with linear regression models residual deviations and applies a second-stage smoothing/thresholding to expose structural changes, so the raw signal needs to be visible before any fitting step hides that structure behind model output.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```





```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```





```{r}
# Load example datasets bundled with harbinger
data(examples_changepoints)
```



```{r}
# Select a simple synthetic time series with labeled change-points
dataset <- examples_changepoints$simple
head(dataset)
```







### Interpret the Result Visually

This first visual pass establishes what the method should react to in the raw series. Keep the method summary in mind here, because changeFinder with linear regression: ChangeFinder with linear regression models residual deviations and applies a second-stage smoothing/thresholding to expose structural changes and the plot tells you whether that structure is clean, weak, local, repeated, or mixed with other effects.

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```







### Configure the Method

The choices below turn the central modeling idea into concrete parameters. They matter because changeFinder with linear regression: ChangeFinder with linear regression models residual deviations and applies a second-stage smoothing/thresholding to expose structural changes, so each argument controls how strongly the method will emphasize that pattern when it later produces change-point candidates.

```{r}
# Define Conformal Forecasting (Linear Regression) change-point model
# - sw_size controls the sliding window length
  model <- hcp_cf_lr(sw_size = 10)
```



```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```







### Run the Core Analysis

This is the moment where the notebook tests its central assumption on actual data. After applying `hcp_cf_lr`, the important question is whether the resulting change-point candidates really correspond to the pattern implied by the method description above, rather than to arbitrary numerical variation.

```{r}
# Detect change-points
  detection <- detect(model, dataset$serie)
```



```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```







### Evaluate What Was Found

The evaluation asks whether the change-point candidates produced by `hcp_cf_lr` match the labeled structure on this dataset. Read the scores as evidence about the method's assumptions in practice, not as detached summary numbers.

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```







### Interpret the Result Visually

This visual check puts the model output back on top of the original signal. What matters now is whether the highlighted change-point candidates line up with the structure suggested by the method, which is the real semantic test of whether the example is teaching the right lesson.

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
  
```



```{r}
# Plot residual scores and threshold
  har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

## References

- Takeuchi, J., Yamanishi, K. (2006). A unifying framework for detecting outliers and change points from time series. IEEE TKDE. doi:10.1109/TKDE.2006.1599387