## Objective

Waypoint combines an autoencoder with a bilateral CUSUM supervisor to detect
regime changes. This notebook keeps the same experimental line as the two
previous Waypoint examples and changes only the encoder architecture to a
convolutional autoencoder.

- Load and visualize the same example change-point dataset used in the previous
  Waypoint notebooks
- Configure `hcp_waypoint()` with a convolutional autoencoder
- Inspect detected change points, evaluate them, and study the reconstruction
  error against the learned decision level

## Method at a glance

Waypoint: an autoencoder learns the current regime from sliding windows, the
window reconstruction error is standardized on a recent buffer, and a bilateral
CUSUM validates persistent deviations. After confirmation, the model is
retrained on the new regime.

## Experimental line

This notebook closes the three-architecture comparison:

- `11`: feed-forward autoencoder, generic baseline
- `12`: LSTM autoencoder, sequential bias inside the window
- `13`: convolutional autoencoder, local-pattern bias inside the window

Again, only the encoder architecture changes. This keeps the interpretation of
results tied to representation bias rather than to unrelated hyperparameter
changes.

## What you will do

- reproduce the same Waypoint workflow with a convolutional encoder
- inspect whether local-pattern bias helps on this regime-change example
- read the evaluation outputs and diagnostic plots in the same way as the first
  two notebooks
- close the comparison line started by the feed-forward baseline

### Prepare the Example

The same dataset is reused because the goal is to compare architectures under a
shared experimental protocol, not to change the underlying change-point task.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/harbinger/master/examples/seed.R"))
# Install Harbinger and DALToolboxDP (if needed)
# install.packages("harbinger")
# install.packages("daltoolboxdp")
```

```{r}
# Load required packages
library(daltoolbox)
library(daltoolboxdp)
library(harbinger)
```

```{r}
# Load example change-point datasets
data(examples_changepoints)
```

```{r}
# Select the same dataset used in the AMOC example
dataset <- examples_changepoints$complex
head(dataset)
```

### Interpret the Result Visually

The raw series is unchanged, but the representation bias is not. A
convolutional encoder emphasizes local motifs inside the sliding window, so the
main experimental question is whether that bias helps distinguish the regime
transition from short-scale variation.

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

### Configure the Method

This is the same Waypoint configuration used in the previous two notebooks. The
sole experimental change is `encoderclass = autoenc_conv_ed`, which emphasizes
local patterns in the input window.

```{r}
# Configure Waypoint with a convolutional autoencoder
model <- hcp_waypoint(
  input_size = 12,
  encode_size = 4,
  warmup = 60,
  retrain_size = 30,
  buffer_size = 40,
  k_factor = 0.35,
  h_low = 3.5,
  h_high = 6,
  prob_tau = 0.997,
  epochs_init = 100,
  epochs_retrain = 100,
  encoderclass = autoenc_conv_ed
)
```

```{r}
# Fit the detector
set_example_seed()
model <- fit(model, dataset$serie)
```

### Run the Core Analysis

This run asks whether a local-pattern representation leads to a more useful
reconstruction-error profile for Waypoint's CUSUM supervisor. Since the
decision logic is unchanged, differences relative to the previous notebooks
should be interpreted as architectural effects.

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected change points
print(detection |> dplyr::filter(event == TRUE))
```

### Evaluate What Was Found

The scores below complete the experimental line. Read them relative to the
feed-forward and LSTM variants to decide whether local-pattern bias is better
or worse for this series and this change-point structure.

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

### Interpret the Result Visually

This final plot checks whether the convolutional representation makes the
transition easier or harder to isolate. The interpretation should stay aligned
with the first two notebooks: compare where the detector fires and how the
reconstruction error separates the change regime from the stable segments.

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot reconstruction error and the learned decision level
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "tau"))
```

## References

- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time
  Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
