## Objective

This notebook demonstrates the fuzzy ensemble `har_ensemble_fuzzy()` using a
set of change-point detectors that behave better on the `simple` benchmark than
the previous placeholder combination.

- load and visualize the simple change-point dataset
- configure a fuzzy ensemble with stronger base detectors
- inspect the final detection and the ensemble score plot

## Method at a glance

The fuzzy ensemble aggregates detector outputs as scores, allows temporal
fuzzification around nearby detections, and can apply non-maximum suppression
to keep the most representative event in a local neighborhood.

## Experimental choice

For this example, the ensemble members are chosen because they behave better on
`examples_changepoints$simple`:

- `hcp_scp(sw = 30)`: exact hit on the labeled changepoint in this dataset
- `hcp_chow()`: exact hit on the labeled changepoint in this dataset
- `hcp_cf_arima(sw_size = 10)`: near-hit around the same transition, useful for
  reinforcing the neighborhood under fuzzy aggregation

## What you will do

- inspect a stronger ensemble composition for this benchmark
- fit the ensemble and run detection
- visualize the fused score over the series

```{r}
# Install Harbinger (if needed)
# install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger)
```

```{r}
# Load example change-point datasets
data(examples_changepoints)
```

```{r}
# Select the simple dataset
dataset <- examples_changepoints$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure the fuzzy ensemble with stronger base detectors for this dataset
model <- har_ensemble_fuzzy(
  hcp_scp(sw = 30),
  hcp_chow(),
  hcp_cf_arima(sw_size = 10)
)
```

```{r}
# Fit the ensemble
model <- fit(model, dataset$serie)
```

```{r}
# Run detection with temporal fuzzification and non-maximum suppression
detection <- detect(model, dataset$serie, time_tolerance = 8, use_nms = TRUE)
```

```{r}
# Show detected change points
print(detection[detection$event, ])
```

```{r}
# Plot the ensemble score over the series
har_ensemble_plot(detection, dataset$serie)
```
