## Objective

This notebook focuses on the ensemble plotting functions. It uses the same
stronger fuzzy ensemble configuration as the previous example so the plots are
driven by detectors that behave better on the `simple` benchmark.

- load the simple change-point dataset
- fit the fuzzy ensemble
- inspect both the fused ensemble view and the per-model contribution view

## Method at a glance

`har_ensemble_plot()` summarizes the fused detection score, while
`har_ensemble_plot_models()` reveals how each base detector contributed to that
score around the candidate change region.

```{r}
# Install Harbinger (if needed)
# install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger)
```

```{r}
# Load example change-point datasets
data(examples_changepoints)
```

```{r}
# Select the simple dataset
dataset <- examples_changepoints$simple
```

```{r}
# Configure the same stronger ensemble used in the previous notebook
model <- har_ensemble_fuzzy(
  hcp_scp(sw = 30),
  hcp_chow(),
  hcp_cf_arima(sw_size = 10)
)
```

```{r}
# Fit the ensemble and run detection
model <- fit(model, dataset$serie)
detection <- detect(model, dataset$serie, time_tolerance = 8, use_nms = TRUE)
```

```{r}
# Plot the fused ensemble score
har_ensemble_plot(detection, dataset$serie)
```

```{r}
# Plot the contribution of each base detector
har_ensemble_plot_models(detection, dataset$serie)
```
