---
title: "KSWIN with `hcp_kswin()`"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
library(daltoolbox)
library(harbinger)
```

## Objective

This notebook demonstrates KSWIN change-point detection on a univariate time series. The detector compares an early sample with the most recent observations inside a sliding window and flags a changepoint when the two distributions differ significantly.

## Method at a glance

KSWIN is a window-based sequential detector for distributional change. In Harbinger it is restricted to one numeric series so that the output can be interpreted directly as virtual drift on the observed signal.

## Prepare the Example

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/harbinger/main/examples/seed.R"))
data(examples_changepoints)
dataset <- examples_changepoints$complex
```

## Visualize the Raw Series

```{r}
har_plot(harbinger(), dataset$serie)
```

## Configure the Method

```{r}
model <- hcp_kswin(window_size = 100, stat_size = 30, alpha = 0.005)
set_example_seed()
model <- fit(model, dataset$serie)
```

## Run Detection

```{r}
detection <- detect(model, dataset$serie)
print(detection[detection$event, ])
```

## Evaluate the Result

```{r}
evaluation <- evaluate(har_eval(), detection$event, dataset$event)
print(evaluation$confMatrix)
```

## Plot the Detections

```{r}
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Raab C, Heusinger M, Schleif FM (2020). Reactive Soft Prototype Computing for Concept Drift Streams. Neurocomputing.
- Bifet A, GavaldĂ  R (2007). Learning from time-changing data with adaptive windowing. SIAM International Conference on Data Mining.
