ChangeFinder with GARCH: ChangeFinder with GARCH models conditional variance dynamics; residual-based scores are smoothed and thresholded to flag volatility regime shifts.

ChangeFinder-GARCH models volatility with GARCH and detects change points by smoothing residual deviations. In this tutorial we:

- Load and visualize a change-point dataset
- Configure and run the GARCH-based change detector (`hcp_garch`)
- Inspect detections, evaluate, and plot residuals with thresholds

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example change-point datasets
data(examples_changepoints)
```

```{r}
# Select the simple dataset
dataset <- examples_changepoints$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure GARCH-based ChangeFinder detector
model <- hcp_garch()
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected change points
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
