ChangeFinder with ARIMA: ChangeFinder with ARIMA models residual deviations and applies a second-stage smoothing/thresholding to highlight structural changes. Implementation wraps ARIMA from `forecast` and uses `harutils()` for thresholds.

ChangeFinder-ARIMA detects change points by modeling residual deviations and smoothing them over a sliding window. In this tutorial we will:

- Load and visualize a simple change-point dataset
- Configure the ChangeFinder-ARIMA detector with a window size
- Inspect detections, evaluate, and plot residuals with thresholds

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example change-point datasets
data(examples_changepoints)
```

```{r}
# Select the simple dataset
dataset <- examples_changepoints$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure ChangeFinder-ARIMA (sw_size controls smoothing window)
model <- hcp_cf_arima(sw_size = 10)
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected change points
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot detections vs. ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot residual magnitude and decision thresholds
har_plot(model, attr(detection, "res"), detection, dataset$event, yline = attr(detection, "threshold"))
```

References 

- Takeuchi, J., Yamanishi, K. (2006). A unifying framework for detecting outliers and change points from time series. IEEE TKDE. doi:10.1109/TKDE.2006.1599387
