---
title: "Online Session with Simulated Source"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
library(daltoolbox)
library(harbinger)
```

## Objective

This notebook validates the minimum end-to-end online workflow in `harbinger`
using the installed package interface.

## Method at a glance

The workflow combines:

- a simulated source
- an online session
- a batch execution strategy
- a memory policy
- trace-based evaluation with Detection Probability and Detection Lag

## Prepare the Example

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/harbinger/master/examples/seed.R"))

serie <- c(1, 1, 1, 1, 5, 6, 7, 8, 20, 21, 22)
reference <- c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE)
source_obj <- har_source_simulated(serie)
```

## Configure the Online Session

```{r}
session <- har_online_session(
  source = source_obj,
  detector = hcp_page_hinkley(min_instances = 3, threshold = 1),
  executor = har_online_refit_full(),
  warmup_size = 3,
  batch_size = 2,
  memory = har_memory_full()
)
```

## Run the Online Workflow

```{r}
session <- fit(session)
session <- run_online(session)
```

## Inspect Final Detection

```{r}
detection <- collect_detection(session)
print(detection)
```

## Inspect the Trace

```{r}
trace <- collect_trace(session)
print(trace)
```

## Inspect the Batch Log

```{r}
batch_log <- collect_batch_log(session)
print(batch_log)
```

## Evaluate the Stream Behavior

```{r}
stream_metrics <- evaluate(har_stream_eval(), trace, reference = reference)
print(stream_metrics$summary)
print(stream_metrics$hard_metrics$confMatrix)
```

