## Objective

The goal of this notebook is to present a complete Harbinger workflow: loading an example dataset, configuring the method used in the notebook, running the analysis, and interpreting the resulting outputs and plots.

## Method at a glance

This notebook showcases motif discovery (repeated subsequences) using Harbinger’s unified interface and base plotting. Across synthetic and ECG datasets, we fit a detector, run discovery, and plot results. The aim is to build intuition for how motifs (and related discords) appear in time series and how Matrix Profile/SAX-based methods surface them.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load motif example datasets and create a base object
data(examples_motifs)
model <- harbinger()
```

```{r}
# Simple synthetic motif dataset
dataset <- examples_motifs$simple
model <- fit(model, dataset$serie)
detection <- detect(model, dataset$serie)
har_plot(model, dataset$serie, detection, dataset$event)

```

```{r}
# ECG sample: MIT-BIH record 100
dataset <- examples_motifs$mitdb100
model <- fit(model, dataset$serie)
detection <- detect(model, dataset$serie)
har_plot(model, dataset$serie, detection, dataset$event)

```

```{r}
# ECG sample: MIT-BIH record 102
dataset <- examples_motifs$mitdb102
model <- fit(model, dataset$serie)
detection <- detect(model, dataset$serie)
har_plot(model, dataset$serie, detection, dataset$event)

```

## References

- Yeh, C.-C. M., et al. (2016). Matrix Profile I/II: All-pairs similarity joins and scalable time series motif/discord discovery. IEEE ICDM.
- Tavenard, R., et al. (2020). tsmp: The Matrix Profile in R. The R Journal. doi:10.32614/RJ-2020-021
- Lin, J., Keogh, E., Lonardi, S., Chiu, B. (2007). A symbolic representation of time series, with implications for streaming algorithms. DMKD, 15, 107–144.
