## Objective

The goal of this notebook is to present a complete Harbinger workflow: loading an example dataset, configuring the method used in the notebook, running the analysis, and interpreting the resulting outputs and plots.

## Method at a glance

SoftED provides soft evaluation by matching detections to ground truth within tolerance windows and assigning partial credit. This notebook shows how to compute both hard and soft metrics and when soft evaluation is more appropriate (e.g., temporal uncertainty around event onset).

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure a simple MLP regressor-based anomaly detector
model <- hanr_ml(ts_mlp(ts_norm_gminmax(), input_size = 5, size = 3, decay = 0))
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection 
detection <- detect(model, dataset$serie)
```

```{r}
# Inspect detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Hard evaluation
evaluation <- evaluate(model, detection$event, dataset$event, evaluation = har_eval())
print(evaluation$confMatrix)
```

```{r}
# Soft evaluation (SoftED) with tolerance window sw_size = 5
result <- evaluate(model, detection$event, dataset$event, evaluation = har_eval_soft(sw_size = 5))
print(result$confMatrix)
```

```{r}
# Evaluation can also be performed directly without a model object
result <- evaluate(har_eval_soft(sw_size = 5), detection$event, dataset$event)
print(result$confMatrix)
```

## References

- Salles, R., Lima, J., Reis, M., Coutinho, R., Pacitti, E., Masseglia, F., Akbarinia, R., Chen, C., Garibaldi, J., Porto, F., Ogasawara, E. (2024). SoftED: Metrics for soft evaluation of time series event detection. Computers and Industrial Engineering. doi:10.1016/j.cie.2024.110728
