Overview and objectives: This notebook demonstrates hard evaluation of event detection results using confusion matrix–based metrics (accuracy, precision, recall, F1). It shows how to compute metrics with `har_eval()` from detections and ground truth, and how to visualize results alongside the series.

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example anomaly datasets
data(examples_anomalies)
```

```{r}
# Select a simple anomaly dataset
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure a simple MLP regressor-based anomaly detector
model <- hanr_ml(ts_mlp(ts_norm_gminmax(), input_size = 5, size = 3, decay = 0))
```

```{r}
# Fit the detector
model <- fit(model, dataset$serie)
```

```{r}
# Run detection
detection <- detect(model, dataset$serie)
```

```{r}
# Inspect detected anomaly indices
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate using hard metrics
evaluation <- evaluate(har_eval(), detection$event, dataset$event)
print(evaluation$confMatrix)
```

References

- Salles, R., Lima, J., Reis, M., Coutinho, R., Pacitti, E., Masseglia, F., Akbarinia, R., Chen, C., Garibaldi, J., Porto, F., Ogasawara, E. (2024). SoftED: Metrics for soft evaluation of time series event detection. Computers and Industrial Engineering. doi:10.1016/j.cie.2024.110728
