## Objective

This notebook introduces the Yahoo Webscope S5 benchmark datasets available in Harbinger. For each dataset object, the goal is to load the full collection with `loadfulldata()`, count the available series, identify the data as univariate or multivariate, and plot the first available signal with `har_plot()`.

## Method at a glance

This is a dataset-orientation notebook. It does not fit a detector. Instead, it shows how to inspect the dataset structure before choosing a modeling strategy. To keep the figures didactic, each plot is capped at the first 1000 observations of the selected series.

## What you will do

- load the packaged mini object and replace it with the full dataset
- count how many series are available in each benchmark object
- identify whether the collection is univariate or multivariate
- plot a short preview of the first available signal and its event labels








### Define the Support Structures

The helper code below standardizes how each collection is counted, typed, and previewed. In a dataset notebook this is not scaffolding for its own sake; it is the mechanism that makes different collections comparable before any detector is discussed.

```{r}
library(harbinger)

dataset_summary <- function(x) {
  first_series <- x[[1]]
  meta_cols <- c("idx", "event", "type", "seq", "seqlen")
  signal_cols <- setdiff(names(first_series), meta_cols)
  dataset_type <- if ("value" %in% names(first_series) || length(signal_cols) == 1) {
    "univariate"
  } else {
    "multivariate"
  }
  plot_column <- if ("value" %in% names(first_series)) "value" else signal_cols[1]

  list(
    n_series = length(x),
    dataset_type = dataset_type,
    signal_cols = signal_cols,
    plot_column = plot_column,
    preview_size = min(1000, nrow(first_series)),
    first_series = first_series
  )
}

show_dataset <- function(x, name) {
  info <- dataset_summary(x)
  cat("Dataset:", name, "\n")
  cat("Number of series:", info$n_series, "\n")
  cat("Dataset type:", info$dataset_type, "\n")
  cat("Signals in the first series:", paste(info$signal_cols, collapse = ", "), "\n")
  cat("Column plotted with har_plot():", info$plot_column, "\n")
  cat("Plot preview length:", info$preview_size, "observations\n")
  invisible(info)
}

plot_dataset_preview <- function(info) {
  preview <- info$first_series[seq_len(info$preview_size), , drop = FALSE]
  har_plot(
    harbinger(),
    preview[[info$plot_column]],
    event = preview$event
  )
}
```

### A1Benchmark
### A1Benchmark

`A1Benchmark` contains real-world labeled univariate series from Yahoo Webscope S5.






This subsection previews A1Benchmark with the same summary routine defined above. The goal is to understand what kind of signal, dimensionality, and labeling this specific collection brings to the benchmark before any method is chosen.

```{r}
data(A1Benchmark)
A1Benchmark <- loadfulldata(A1Benchmark)
a1_info <- show_dataset(A1Benchmark, "A1Benchmark")
```





```{r}
plot_dataset_preview(a1_info)
```

### A2Benchmark
### A2Benchmark

`A2Benchmark` contains synthetic labeled univariate series designed for anomaly-detection benchmarking.






This subsection previews A2Benchmark with the same summary routine defined above. The goal is to understand what kind of signal, dimensionality, and labeling this specific collection brings to the benchmark before any method is chosen.

```{r}
data(A2Benchmark)
A2Benchmark <- loadfulldata(A2Benchmark)
a2_info <- show_dataset(A2Benchmark, "A2Benchmark")
```





```{r}
plot_dataset_preview(a2_info)
```

### A3Benchmark
### A3Benchmark

`A3Benchmark` contains synthetic Yahoo series with labeled anomalous observations.






This subsection previews A3Benchmark with the same summary routine defined above. The goal is to understand what kind of signal, dimensionality, and labeling this specific collection brings to the benchmark before any method is chosen.

```{r}
data(A3Benchmark)
A3Benchmark <- loadfulldata(A3Benchmark)
a3_info <- show_dataset(A3Benchmark, "A3Benchmark")
```





```{r}
plot_dataset_preview(a3_info)
```

### A4Benchmark
### A4Benchmark

`A4Benchmark` combines synthetic anomalies and structural changes in the Yahoo benchmark family.






This subsection previews A4Benchmark with the same summary routine defined above. The goal is to understand what kind of signal, dimensionality, and labeling this specific collection brings to the benchmark before any method is chosen.

```{r}
data(A4Benchmark)
A4Benchmark <- loadfulldata(A4Benchmark)
a4_info <- show_dataset(A4Benchmark, "A4Benchmark")
```





```{r}
plot_dataset_preview(a4_info)
```

## References

- Yahoo Webscope S5 benchmark documentation and downstream studies on labeled anomaly detection.
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3