## Objective

This notebook introduces the Numenta Anomaly Benchmark datasets exposed by Harbinger. For each NAB dataset object, it demonstrates how to load the full collection with `loadfulldata()`, count the available series, inspect the structure, and plot the first available signal with `har_plot()`.

## Method at a glance

The focus is on dataset inspection. NAB is often used as a benchmark family, so this notebook helps the reader understand the shape of each object before choosing a detection workflow. To avoid overly long figures, each plot uses at most the first 1000 observations of the first series.

## What you will do

- load each NAB object with `data()` and `loadfulldata()`
- count how many series are available in the full dataset
- verify that the first series is univariate
- plot a short preview of the first signal and its labels

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
library(harbinger)

dataset_summary <- function(x) {
  first_series <- x[[1]]
  meta_cols <- c("idx", "event", "type", "seq", "seqlen")
  signal_cols <- setdiff(names(first_series), meta_cols)
  dataset_type <- if ("value" %in% names(first_series) || length(signal_cols) == 1) "univariate" else "multivariate"
  plot_column <- if ("value" %in% names(first_series)) "value" else signal_cols[1]

  list(
    n_series = length(x),
    dataset_type = dataset_type,
    signal_cols = signal_cols,
    plot_column = plot_column,
    preview_size = min(1000, nrow(first_series)),
    first_series = first_series
  )
}

show_dataset <- function(x, name) {
  info <- dataset_summary(x)
  cat("Dataset:", name, "\n")
  cat("Number of series:", info$n_series, "\n")
  cat("Dataset type:", info$dataset_type, "\n")
  cat("Signals in the first series:", paste(info$signal_cols, collapse = ", "), "\n")
  cat("Column plotted with har_plot():", info$plot_column, "\n")
  cat("Plot preview length:", info$preview_size, "observations\n")
  invisible(info)
}

plot_dataset_preview <- function(info) {
  preview <- info$first_series[seq_len(info$preview_size), , drop = FALSE]
  har_plot(
    harbinger(),
    preview[[info$plot_column]],
    event = preview$event
  )
}
```

### nab_artificialWithAnomaly

```{r}
data(nab_artificialWithAnomaly)
nab_artificialWithAnomaly <- loadfulldata(nab_artificialWithAnomaly)
nab_artificial_info <- show_dataset(nab_artificialWithAnomaly, "nab_artificialWithAnomaly")
```

```{r}
plot_dataset_preview(nab_artificial_info)
```

### nab_realAdExchange

```{r}
data(nab_realAdExchange)
nab_realAdExchange <- loadfulldata(nab_realAdExchange)
nab_adexchange_info <- show_dataset(nab_realAdExchange, "nab_realAdExchange")
```

```{r}
plot_dataset_preview(nab_adexchange_info)
```

### nab_realAWSCloudwatch

```{r}
data(nab_realAWSCloudwatch)
nab_realAWSCloudwatch <- loadfulldata(nab_realAWSCloudwatch)
nab_aws_info <- show_dataset(nab_realAWSCloudwatch, "nab_realAWSCloudwatch")
```

```{r}
plot_dataset_preview(nab_aws_info)
```

### nab_realKnownCause

```{r}
data(nab_realKnownCause)
nab_realKnownCause <- loadfulldata(nab_realKnownCause)
nab_known_info <- show_dataset(nab_realKnownCause, "nab_realKnownCause")
```

```{r}
plot_dataset_preview(nab_known_info)
```

### nab_realTraffic

```{r}
data(nab_realTraffic)
nab_realTraffic <- loadfulldata(nab_realTraffic)
nab_traffic_info <- show_dataset(nab_realTraffic, "nab_realTraffic")
```

```{r}
plot_dataset_preview(nab_traffic_info)
```

### nab_realTweets

```{r}
data(nab_realTweets)
nab_realTweets <- loadfulldata(nab_realTweets)
nab_tweets_info <- show_dataset(nab_realTweets, "nab_realTweets")
```

```{r}
plot_dataset_preview(nab_tweets_info)
```

## References

- Lavin, A., Ahmad, S. (2015). Evaluating real-time anomaly detection algorithms: the Numenta Anomaly Benchmark.
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3

