## Objective

This notebook introduces the multivariate 3W oil-well datasets bundled through Harbinger. For each dataset object, it loads the full collection with `loadfulldata()`, reports how many series are available, confirms that the collection is multivariate, and plots the first sensor column from the first series with `har_plot()`.

## Method at a glance

This is a data-inspection notebook aimed at multivariate industrial series. The key point is to show how to identify the sensor columns and choose a first variable for visual inspection. To keep the plots readable, each preview uses at most the first 1000 observations.

## What you will do

- load each 3W dataset object and expand it to the full collection
- count the number of multivariate series available
- inspect the first-series sensor columns
- plot a short preview of the first sensor column together with its labels

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
library(harbinger)

dataset_summary <- function(x) {
  first_series <- x[[1]]
  meta_cols <- c("idx", "event", "type", "seq", "seqlen")
  signal_cols <- setdiff(names(first_series), meta_cols)
  dataset_type <- if ("value" %in% names(first_series) || length(signal_cols) == 1) "univariate" else "multivariate"
  plot_column <- if ("value" %in% names(first_series)) "value" else signal_cols[1]

  list(
    n_series = length(x),
    dataset_type = dataset_type,
    signal_cols = signal_cols,
    plot_column = plot_column,
    preview_size = min(1000, nrow(first_series)),
    first_series = first_series
  )
}

show_dataset <- function(x, name) {
  info <- dataset_summary(x)
  cat("Dataset:", name, "\n")
  cat("Number of series:", info$n_series, "\n")
  cat("Dataset type:", info$dataset_type, "\n")
  cat("Signals in the first series:", paste(info$signal_cols, collapse = ", "), "\n")
  cat("Column plotted with har_plot():", info$plot_column, "\n")
  cat("Plot preview length:", info$preview_size, "observations\n")
  invisible(info)
}

plot_dataset_preview <- function(info) {
  preview <- info$first_series[seq_len(info$preview_size), , drop = FALSE]
  har_plot(
    harbinger(),
    preview[[info$plot_column]],
    event = preview$event
  )
}
```

### oil_3w_Type_1

```{r}
data(oil_3w_Type_1)
oil_3w_Type_1 <- loadfulldata(oil_3w_Type_1)
oil_type_1_info <- show_dataset(oil_3w_Type_1, "oil_3w_Type_1")
```

```{r}
plot_dataset_preview(oil_type_1_info)
```

### oil_3w_Type_2

```{r}
data(oil_3w_Type_2)
oil_3w_Type_2 <- loadfulldata(oil_3w_Type_2)
oil_type_2_info <- show_dataset(oil_3w_Type_2, "oil_3w_Type_2")
```

```{r}
plot_dataset_preview(oil_type_2_info)
```

### oil_3w_Type_4

```{r}
data(oil_3w_Type_4)
oil_3w_Type_4 <- loadfulldata(oil_3w_Type_4)
oil_type_4_info <- show_dataset(oil_3w_Type_4, "oil_3w_Type_4")
```

```{r}
plot_dataset_preview(oil_type_4_info)
```

### oil_3w_Type_5

```{r}
data(oil_3w_Type_5)
oil_3w_Type_5 <- loadfulldata(oil_3w_Type_5)
oil_type_5_info <- show_dataset(oil_3w_Type_5, "oil_3w_Type_5")
```

```{r}
plot_dataset_preview(oil_type_5_info)
```

### oil_3w_Type_6

```{r}
data(oil_3w_Type_6)
oil_3w_Type_6 <- loadfulldata(oil_3w_Type_6)
oil_type_6_info <- show_dataset(oil_3w_Type_6, "oil_3w_Type_6")
```

```{r}
plot_dataset_preview(oil_type_6_info)
```

### oil_3w_Type_7

```{r}
data(oil_3w_Type_7)
oil_3w_Type_7 <- loadfulldata(oil_3w_Type_7)
oil_type_7_info <- show_dataset(oil_3w_Type_7, "oil_3w_Type_7")
```

```{r}
plot_dataset_preview(oil_type_7_info)
```

### oil_3w_Type_8

```{r}
data(oil_3w_Type_8)
oil_3w_Type_8 <- loadfulldata(oil_3w_Type_8)
oil_type_8_info <- show_dataset(oil_3w_Type_8, "oil_3w_Type_8")
```

```{r}
plot_dataset_preview(oil_type_8_info)
```

## References

- 3W dataset documentation from the UCI Machine Learning Repository.
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3

