## Objective

This notebook shows how to load the full `gecco` dataset with `loadfulldata()`, inspect how many series are available, verify whether the collection is univariate or multivariate, and plot the first signal with `har_plot()`.

## Method at a glance

The purpose here is exploratory orientation. Before modeling, the reader can confirm how the dataset object is organized and which column should be used as the first signal in a quick inspection. To keep the visualization readable, the plot shows at most the first 1000 observations of the first series.

## What you will do

- load the bundled `gecco` object and replace it with the full dataset
- count the available series in the collection
- inspect the signal columns present in the first series
- plot a didactic preview of the first available signal with its event labels








### Define the Support Structures

The helper code below standardizes how each collection is counted, typed, and previewed. In a dataset notebook this is not scaffolding for its own sake; it is the mechanism that makes different collections comparable before any detector is discussed.

```{r}
library(harbinger)

dataset_summary <- function(x) {
  first_series <- x[[1]]
  meta_cols <- c("idx", "event", "type", "seq", "seqlen")
  signal_cols <- setdiff(names(first_series), meta_cols)
  dataset_type <- if ("value" %in% names(first_series) || length(signal_cols) == 1) "univariate" else "multivariate"
  plot_column <- if ("value" %in% names(first_series)) "value" else signal_cols[1]

  list(
    n_series = length(x),
    dataset_type = dataset_type,
    signal_cols = signal_cols,
    plot_column = plot_column,
    preview_size = min(1000, nrow(first_series)),
    first_series = first_series
  )
}

show_dataset <- function(x, name) {
  info <- dataset_summary(x)
  cat("Dataset:", name, "\n")
  cat("Number of series:", info$n_series, "\n")
  cat("Dataset type:", info$dataset_type, "\n")
  cat("Signals in the first series:", paste(info$signal_cols, collapse = ", "), "\n")
  cat("Column plotted with har_plot():", info$plot_column, "\n")
  cat("Plot preview length:", info$preview_size, "observations\n")
  invisible(info)
}

plot_dataset_preview <- function(info) {
  preview <- info$first_series[seq_len(info$preview_size), , drop = FALSE]
  har_plot(
    harbinger(),
    preview[[info$plot_column]],
    event = preview$event
  )
}
```







### Prepare the Example

This setup loads the collections used in the notebook and keeps the dataset description concrete from the start. In this case the analysis begins before modeling, because the reader needs to see how the collection is organized before deciding what kind of method makes sense.

```{r}
data(gecco)
gecco <- loadfulldata(gecco)
gecco_info <- show_dataset(gecco, "gecco")
```







### Interpret the Result Visually

This first visual pass establishes what the method should react to in the raw series. Keep the method summary in mind here, because the purpose here is exploratory orientation and the plot tells you whether that structure is clean, weak, local, repeated, or mixed with other effects.

```{r}
plot_dataset_preview(gecco_info)
```

## References

- GECCO Challenge 2018 material for water-quality event detection.
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3