## Objective

This notebook shows how to load the full `gecco` dataset with `loadfulldata()`, inspect how many series are available, verify whether the collection is univariate or multivariate, and plot the first signal with `har_plot()`.

## Method at a glance

The purpose here is exploratory orientation. Before modeling, the reader can confirm how the dataset object is organized and which column should be used as the first signal in a quick inspection. To keep the visualization readable, the plot shows at most the first 1000 observations of the first series.

## What you will do

- load the bundled `gecco` object and replace it with the full dataset
- count the available series in the collection
- inspect the signal columns present in the first series
- plot a didactic preview of the first available signal with its event labels

## Walkthrough

```{r}
library(harbinger)

dataset_summary <- function(x) {
  first_series <- x[[1]]
  meta_cols <- c("idx", "event", "type", "seq", "seqlen")
  signal_cols <- setdiff(names(first_series), meta_cols)
  dataset_type <- if ("value" %in% names(first_series) || length(signal_cols) == 1) "univariate" else "multivariate"
  plot_column <- if ("value" %in% names(first_series)) "value" else signal_cols[1]

  list(
    n_series = length(x),
    dataset_type = dataset_type,
    signal_cols = signal_cols,
    plot_column = plot_column,
    preview_size = min(1000, nrow(first_series)),
    first_series = first_series
  )
}

show_dataset <- function(x, name) {
  info <- dataset_summary(x)
  cat("Dataset:", name, "\n")
  cat("Number of series:", info$n_series, "\n")
  cat("Dataset type:", info$dataset_type, "\n")
  cat("Signals in the first series:", paste(info$signal_cols, collapse = ", "), "\n")
  cat("Column plotted with har_plot():", info$plot_column, "\n")
  cat("Plot preview length:", info$preview_size, "observations\n")
  invisible(info)
}

plot_dataset_preview <- function(info) {
  preview <- info$first_series[seq_len(info$preview_size), , drop = FALSE]
  har_plot(
    harbinger(),
    preview[[info$plot_column]],
    event = preview$event
  )
}
```

```{r}
data(gecco)
gecco <- loadfulldata(gecco)
gecco_info <- show_dataset(gecco, "gecco")
```

```{r}
plot_dataset_preview(gecco_info)
```

## References

- GECCO Challenge 2018 material for water-quality event detection.
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
