## Objective

This notebook introduces the Numenta Anomaly Benchmark datasets exposed by Harbinger. For each NAB dataset object, it demonstrates how to load the full collection with `loadfulldata()`, count the available series, inspect the structure, and plot the first available signal with `har_plot()`.

## Method at a glance

The focus is on dataset inspection. NAB is often used as a benchmark family, so this notebook helps the reader understand the shape of each object before choosing a detection workflow.

## What you will do

- load each NAB object with `data()` and `loadfulldata()`
- count how many series are available in the full dataset
- verify that the first series is univariate
- plot the first signal and its labels

## Walkthrough

```{r}
library(harbinger)

dataset_summary <- function(x) {
  first_series <- x[[1]]
  meta_cols <- c("idx", "event", "type", "seq", "seqlen")
  signal_cols <- setdiff(names(first_series), meta_cols)
  dataset_type <- if ("value" %in% names(first_series) || length(signal_cols) == 1) "univariate" else "multivariate"
  plot_column <- if ("value" %in% names(first_series)) "value" else signal_cols[1]

  list(
    n_series = length(x),
    dataset_type = dataset_type,
    signal_cols = signal_cols,
    plot_column = plot_column,
    first_series = first_series
  )
}

show_dataset <- function(x, name) {
  info <- dataset_summary(x)
  cat("Dataset:", name, "\n")
  cat("Number of series:", info$n_series, "\n")
  cat("Dataset type:", info$dataset_type, "\n")
  cat("Signals in the first series:", paste(info$signal_cols, collapse = ", "), "\n")
  cat("Column plotted with har_plot():", info$plot_column, "\n")
  invisible(info)
}
```

### nab_artificialWithAnomaly

```{r}
data(nab_artificialWithAnomaly)
nab_artificialWithAnomaly <- loadfulldata(nab_artificialWithAnomaly)
nab_artificial_info <- show_dataset(nab_artificialWithAnomaly, "nab_artificialWithAnomaly")
```

```{r}
har_plot(
  harbinger(),
  nab_artificial_info$first_series[[nab_artificial_info$plot_column]],
  event = nab_artificial_info$first_series$event
)
```

### nab_realAdExchange

```{r}
data(nab_realAdExchange)
nab_realAdExchange <- loadfulldata(nab_realAdExchange)
nab_adexchange_info <- show_dataset(nab_realAdExchange, "nab_realAdExchange")
```

```{r}
har_plot(
  harbinger(),
  nab_adexchange_info$first_series[[nab_adexchange_info$plot_column]],
  event = nab_adexchange_info$first_series$event
)
```

### nab_realAWSCloudwatch

```{r}
data(nab_realAWSCloudwatch)
nab_realAWSCloudwatch <- loadfulldata(nab_realAWSCloudwatch)
nab_aws_info <- show_dataset(nab_realAWSCloudwatch, "nab_realAWSCloudwatch")
```

```{r}
har_plot(
  harbinger(),
  nab_aws_info$first_series[[nab_aws_info$plot_column]],
  event = nab_aws_info$first_series$event
)
```

### nab_realKnownCause

```{r}
data(nab_realKnownCause)
nab_realKnownCause <- loadfulldata(nab_realKnownCause)
nab_known_info <- show_dataset(nab_realKnownCause, "nab_realKnownCause")
```

```{r}
har_plot(
  harbinger(),
  nab_known_info$first_series[[nab_known_info$plot_column]],
  event = nab_known_info$first_series$event
)
```

### nab_realTraffic

```{r}
data(nab_realTraffic)
nab_realTraffic <- loadfulldata(nab_realTraffic)
nab_traffic_info <- show_dataset(nab_realTraffic, "nab_realTraffic")
```

```{r}
har_plot(
  harbinger(),
  nab_traffic_info$first_series[[nab_traffic_info$plot_column]],
  event = nab_traffic_info$first_series$event
)
```

### nab_realTweets

```{r}
data(nab_realTweets)
nab_realTweets <- loadfulldata(nab_realTweets)
nab_tweets_info <- show_dataset(nab_realTweets, "nab_realTweets")
```

```{r}
har_plot(
  harbinger(),
  nab_tweets_info$first_series[[nab_tweets_info$plot_column]],
  event = nab_tweets_info$first_series$event
)
```

## References

- Lavin, A., Ahmad, S. (2015). Evaluating real-time anomaly detection algorithms: the Numenta Anomaly Benchmark.
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
