## Objective

This notebook demonstrates how to inspect the UCR Anomaly Archive datasets distributed by Harbinger. Each example loads the full dataset with `loadfulldata()`, counts the available series, identifies the structure, and plots the first signal with `har_plot()`.

## Method at a glance

The emphasis is on understanding the data layout before applying an event-detection method. The same inspection pattern is repeated for each UCR dataset object.

## What you will do

- load each UCR dataset object and expand it to the full collection
- count the number of series available
- inspect the first series and confirm the signal column
- plot the first signal and its labels

## Walkthrough

```{r}
library(harbinger)

dataset_summary <- function(x) {
  first_series <- x[[1]]
  meta_cols <- c("idx", "event", "type", "seq", "seqlen")
  signal_cols <- setdiff(names(first_series), meta_cols)
  dataset_type <- if ("value" %in% names(first_series) || length(signal_cols) == 1) "univariate" else "multivariate"
  plot_column <- if ("value" %in% names(first_series)) "value" else signal_cols[1]

  list(
    n_series = length(x),
    dataset_type = dataset_type,
    signal_cols = signal_cols,
    plot_column = plot_column,
    first_series = first_series
  )
}

show_dataset <- function(x, name) {
  info <- dataset_summary(x)
  cat("Dataset:", name, "\n")
  cat("Number of series:", info$n_series, "\n")
  cat("Dataset type:", info$dataset_type, "\n")
  cat("Signals in the first series:", paste(info$signal_cols, collapse = ", "), "\n")
  cat("Column plotted with har_plot():", info$plot_column, "\n")
  invisible(info)
}
```

### ucr_ecg

```{r}
data(ucr_ecg)
ucr_ecg <- loadfulldata(ucr_ecg)
ucr_ecg_info <- show_dataset(ucr_ecg, "ucr_ecg")
```

```{r}
har_plot(
  harbinger(),
  ucr_ecg_info$first_series[[ucr_ecg_info$plot_column]],
  event = ucr_ecg_info$first_series$event
)
```

### ucr_nasa

```{r}
data(ucr_nasa)
ucr_nasa <- loadfulldata(ucr_nasa)
ucr_nasa_info <- show_dataset(ucr_nasa, "ucr_nasa")
```

```{r}
har_plot(
  harbinger(),
  ucr_nasa_info$first_series[[ucr_nasa_info$plot_column]],
  event = ucr_nasa_info$first_series$event
)
```

### ucr_int_bleeding

```{r}
data(ucr_int_bleeding)
ucr_int_bleeding <- loadfulldata(ucr_int_bleeding)
ucr_bleeding_info <- show_dataset(ucr_int_bleeding, "ucr_int_bleeding")
```

```{r}
har_plot(
  harbinger(),
  ucr_bleeding_info$first_series[[ucr_bleeding_info$plot_column]],
  event = ucr_bleeding_info$first_series$event
)
```

### ucr_power_demand

```{r}
data(ucr_power_demand)
ucr_power_demand <- loadfulldata(ucr_power_demand)
ucr_power_info <- show_dataset(ucr_power_demand, "ucr_power_demand")
```

```{r}
har_plot(
  harbinger(),
  ucr_power_info$first_series[[ucr_power_info$plot_column]],
  event = ucr_power_info$first_series$event
)
```

## References

- UCR Time Series Anomaly Archive.
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
