## Objective

SAX (Symbolic Aggregate approXimation) maps z-normalized values to symbols; recurring words indicate motifs. In this example we:

- Load and visualize a motif dataset
- Configure `hmo_sax(a, w, qtd)` and run detection
- Inspect and evaluate motif occurrences

## Method at a glance

SAX motif discovery: SAX (Symbolic Aggregate approXimation) discretizes z-normalized subsequences into symbolic words using breakpoints derived from a Gaussian distribution. Motifs emerge as frequently occurring words or closely related symbolic patterns. Harbinger wraps SAX-based discovery and returns motif locations for plotting and evaluation.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example motif datasets
data(examples_motifs)
```

```{r}
# Select the simple motif dataset
dataset <- examples_motifs$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure SAX-based motif discovery (alphabet=26, word=3, min occurrences=3)
model <- hmo_sax(26, 3, 3)
```

```{r}
# Fit the detector (learns binning thresholds)
model <- fit(model, dataset$serie)
```

```{r}
# Run motif discovery
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected motif starts
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot motifs and ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Lin, J., Keogh, E., Lonardi, S., Chiu, B. (2007). A symbolic representation of time series, with implications for streaming algorithms. Data Mining and Knowledge Discovery, 15, 107–144.
