## Objective

This notebook demonstrates discord (anomaly) discovery using Matrix Profile with the STOMP algorithm via `hdis_mp(mode = "stomp", ...)`. Discords are subsequences maximally dissimilar to the rest. Steps: load packages/data, visualize, define the discord model (subsequence length and count), fit, detect, evaluate, and plot.

## Method at a glance

STOMP motif discovery: Matrix Profile-based discord discovery identifies rare subsequences whose nearest-neighbor distance is large. STOMP approximates the profile via random order calculations for scalability.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_motifs)
```

```{r}
# Select an ECG time series with annotated anomalies
dataset <- examples_motifs$mitdb102
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define Matrix Profile discord model (STOMP)
# - w: subsequence length (window)
# - qtd: number of discords to retrieve
model <- hdis_mp(mode = "stomp", w = 25, qtd = 10)
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect discords
  suppressMessages(detection <- detect(model, dataset$serie))
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
  
```

## References

- Yeh, C.-C. M., et al. (2016). Matrix Profile I/II: All-pairs similarity joins and scalable time series motif/discord discovery. IEEE ICDM.
- Tavenard, R., et al. (2020). tsmp: The Matrix Profile in R. The R Journal. doi:10.32614/RJ-2020-021
