## Objective

SAX-based discord discovery identifies rare, dissimilar subsequences by discretizing the series and finding unique words with high entropy. We will:

- Load and visualize a motif/discord dataset
- Configure `hdis_sax(a, w)` and run discovery
- Inspect and evaluate discord occurrences

## Method at a glance

SAX motif discovery: SAX discretizes z-normalized subsequences into symbolic words; discords emerge as rare words or windows with large symbolic distance to their neighbors.

## What you will do

- understand the purpose of the example and when the technique is useful
- follow the workflow from data loading to model fitting and detection
- inspect the evaluation outputs and the diagnostic plots produced by Harbinger

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example motif/discord datasets
data(examples_motifs)
```

```{r}
# Select an ECG sample (mitdb102)
dataset <- examples_motifs$mitdb102
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure SAX-based discord discovery (alphabet=26, word=25)
model <- hdis_sax(26, 25)
```

```{r}
# Fit the detector (learns binning thresholds)
model <- fit(model, dataset$serie)
```

```{r}
# Run discord discovery
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected discord starts
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot discords and ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Lin, J., Keogh, E., Lonardi, S., Chiu, B. (2007). A symbolic representation of time series, with implications for streaming algorithms. Data Mining and Knowledge Discovery, 15, 107–144.
