SAX motif discovery: SAX discretizes z-normalized subsequences into symbolic words; discords emerge as rare words or windows with large symbolic distance to their neighbors.

SAX-based discord discovery identifies rare, dissimilar subsequences by discretizing the series and finding unique words with high entropy. We will:

- Load and visualize a motif/discord dataset
- Configure `hdis_sax(a, w)` and run discovery
- Inspect and evaluate discord occurrences

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example motif/discord datasets
data(examples_motifs)
```

```{r}
# Select an ECG sample (mitdb102)
dataset <- examples_motifs$mitdb102
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure SAX-based discord discovery (alphabet=26, word=25)
model <- hdis_sax(26, 25)
```

```{r}
# Fit the detector (learns binning thresholds)
model <- fit(model, dataset$serie)
```

```{r}
# Run discord discovery
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected discord starts
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot discords and ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

References 
- Lin, J., Keogh, E., Lonardi, S., Chiu, B. (2007). A symbolic representation of time series, with implications for streaming algorithms. Data Mining and Knowledge Discovery, 15, 107–144.
