XSAX motif discovery: XSAX extends SAX to a larger alphanumeric alphabet, allowing finer symbolic resolution. Motifs are frequent symbolic words or closely related patterns.

XSAX extends SAX to a larger alphanumeric alphabet, enabling finer symbolic resolution for motif discovery. In this tutorial we:

- Load and visualize a motif dataset
- Configure `hmo_xsax(a, w, qtd)` and run detection
- Inspect and evaluate motif occurrences

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example motif datasets
data(examples_motifs)
```

```{r}
# Select the simple motif dataset
dataset <- examples_motifs$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure XSAX motif discovery (alphabet=37, word=3, min occurrences=3)
model <- hmo_xsax(37, 3, 3)
```

```{r}
# Fit the detector (learns binning thresholds)
model <- fit(model, dataset$serie)
```

```{r}
# Run motif discovery
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected motif starts
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot motifs and ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

References 
- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
