SAX motif discovery: SAX (Symbolic Aggregate approXimation) discretizes z-normalized subsequences into symbolic words using breakpoints derived from a Gaussian distribution. Motifs emerge as frequently occurring words or closely related symbolic patterns. Harbinger wraps SAX-based discovery and returns motif locations for plotting and evaluation.

SAX (Symbolic Aggregate approXimation) maps z-normalized values to symbols; recurring words indicate motifs. In this example we:

- Load and visualize a motif dataset
- Configure `hmo_sax(a, w, qtd)` and run detection
- Inspect and evaluate motif occurrences

```{r}
# Install Harbinger (if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example motif datasets
data(examples_motifs)
```

```{r}
# Select the simple motif dataset
dataset <- examples_motifs$simple
head(dataset)
```

```{r}
# Plot the raw time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Configure SAX-based motif discovery (alphabet=26, word=3, min occurrences=3)
model <- hmo_sax(26, 3, 3)
```

```{r}
# Fit the detector (learns binning thresholds)
model <- fit(model, dataset$serie)
```

```{r}
# Run motif discovery
detection <- detect(model, dataset$serie)
```

```{r}
# Show detected motif starts
print(detection |> dplyr::filter(event == TRUE))
```

```{r}
# Evaluate detections against labels
evaluation <- evaluate(model, detection$event, dataset$event)
print(evaluation$confMatrix)
```

```{r}
# Plot motifs and ground truth
har_plot(model, dataset$serie, detection, dataset$event)
```

References 
- Lin, J., Keogh, E., Lonardi, S., Chiu, B. (2007). A symbolic representation of time series, with implications for streaming algorithms. Data Mining and Knowledge Discovery, 15, 107–144.
