SCRIMP motif discovery: Matrix Profile methods compute nearest-neighbor distances for subsequences. SCRIMP incrementally approximates the profile and refines it efficiently. Provided via `tsmp` and wrapped by `hmo_mp()`.

Objectives: This Rmd demonstrates motif discovery using Matrix Profile with SCRIMP via `hmo_mp("scrimp", ...)`. SCRIMP is an incremental algorithm for computing the matrix profile. Steps: load packages/data, visualize, set subsequence length and count, fit, detect, evaluate, and plot.

```{r}
# Install Harbinger (only once, if needed)
#install.packages("harbinger")
```

```{r}
# Load required packages
library(daltoolbox)
library(harbinger) 
```

```{r}
# Load example datasets bundled with harbinger
data(examples_motifs)
```

```{r}
# Select a simple example time series
dataset <- examples_motifs$simple
head(dataset)
```

```{r}
# Plot the time series
har_plot(harbinger(), dataset$serie)
```

```{r}
# Define Matrix Profile (SCRIMP) motif model
# - second arg: subsequence length (window)
# - third arg: number of motifs to retrieve
  model <- hmo_mp("scrimp", 4, 3)
```

```{r}
# Fit the model
  model <- fit(model, dataset$serie)
```

```{r}
# Detect motifs
  detection <- detect(model, dataset$serie)
```

```{r}
# Show only timestamps flagged as events
  print(detection |> dplyr::filter(event==TRUE))
```

```{r}
# Evaluate detections against ground-truth labels
  evaluation <- evaluate(model, detection$event, dataset$event)
  print(evaluation$confMatrix)
```

```{r}
# Plot detections over the series
  har_plot(model, dataset$serie, detection, dataset$event)
  
```

References 
- Yeh, C.-C. M., et al. (2016). Matrix Profile I/II: All-pairs similarity joins and scalable time series motif/discord discovery. IEEE ICDM.
- Tavenard, R., et al. (2020). tsmp: The Matrix Profile in R. The R Journal. doi:10.32614/RJ-2020-021
