## Objective

This tutorial introduces the core Harbinger workflow on a small labeled anomaly series. The goal is to help the reader see the minimum sequence of actions: load data, create a model, detect events, inspect the output, and plot the result.

## Method at a glance

We use the default `harbinger()` model as a first contact with the package interface. The emphasis here is not on a specific detection technique, but on understanding the common workflow shared by many Harbinger objects.

## What you will do

- load a bundled example dataset
- inspect the structure of the series and event labels
- create the default Harbinger model
- run detection
- visualize the detected events

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
# Install Harbinger (if needed)
# install.packages("harbinger")
```

```{r}
# Load the package
library(harbinger)
```

```{r}
# Load a small labeled anomaly dataset distributed with the package
data(examples_anomalies)
dataset <- examples_anomalies$simple
head(dataset)
```

```{r}
# Plot the series with the known labeled events
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

```{r}
# Create the default Harbinger detector
model <- harbinger()
```

```{r}
# Run detection on the series
detection <- detect(model, dataset$serie)
head(detection)
```

```{r}
# Plot detections against the labeled events
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3

