## Tutorial 01 - First Steps

The first objective is simple: understand the basic Harbinger workflow on a labeled series. In the examples that follow, this same rhythm will appear repeatedly: load data, configure a method, run the analysis, inspect the output, and plot the result.

This notebook uses the default `harbinger()` object because the focus is not on a specific detector yet. The goal is to make the package interface familiar before introducing specialized methods.

Conceptually, this object represents the common contract of the package: a Harbinger method receives a series and returns a standardized event-oriented output that can later be plotted or evaluated. Learning that contract first makes the later detectors much easier to understand.

```{r}
# Install Harbinger (if needed)
# install.packages("harbinger")
```

Load the package and a small labeled anomaly series.
```{r}
library(harbinger)

data(examples_anomalies)
dataset <- examples_anomalies$simple
head(dataset)
```

Before running a method, it helps to see the series and the known event labels.
```{r}
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

Now create the default Harbinger model.
```{r}
model <- harbinger()
```

Run the analysis and inspect the returned structure.
```{r}
detection <- detect(model, dataset$serie)
head(detection)
```

Finish by plotting the detections against the labeled events.
```{r}
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
