## Tutorial 03 - Plotting Series

`har_plot()` is part of the analytical workflow, not only a presentation helper. A quick plot often reveals trends, level shifts, bursts, and labeled events before any detector is configured.

This tutorial shows the same plotting interface on both a univariate and a multivariate example.

The underlying idea is simple but important: in time-series mining, visual inspection is often the first diagnostic technique. It helps the reader connect the raw signal with later decisions about windows, thresholds, transformations, or detector families.

```{r}
library(harbinger)
```

Start with a univariate anomaly series.
```{r}
data(examples_anomalies)
uni <- examples_anomalies$simple
har_plot(harbinger(), uni$serie, event = uni$event)
```

Now load a multivariate dataset and inspect which signal columns are available.
```{r}
data(oil_3w_Type_1)
oil_3w_Type_1 <- loadfulldata(oil_3w_Type_1)
first_mv <- oil_3w_Type_1[[1]]
meta_cols <- c("idx", "event", "type", "seq", "seqlen")
signal_cols <- setdiff(names(first_mv), meta_cols)
signal_cols
```

Plot one signal column from the multivariate series with the same interface.
```{r}
har_plot(
  harbinger(),
  first_mv[[signal_cols[1]]],
  event = first_mv$event
)
```

## References

- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3