## Objective

This tutorial introduces a model-based anomaly detector using `hanr_arima()`. The goal is to show how Harbinger can detect unusual behavior by modeling expected dynamics and then analyzing residual deviations.

## Method at a glance

`hanr_arima()` fits an ARIMA model to the series, computes residual magnitudes, and flags unusually large residuals as anomalies. This moves from a simple distribution-based baseline to a temporal model that tries to explain normal behavior first.

## What you will do

- fit an ARIMA-based detector
- run anomaly detection on a labeled series
- inspect the evaluation output
- plot residual magnitudes and thresholds

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough







### Prepare the Example

We begin by organizing the environment, loading the packages, and selecting the dataset used in the notebook. This part is intentionally more direct: the goal is to make the starting point explicit before the method-specific reasoning begins.

```{r}
library(daltoolbox)
library(harbinger)
```



```{r}
data(examples_anomalies)
dataset <- examples_anomalies$simple
```







### Configure the Method

The next step is to instantiate the method and, when necessary, fit it to the selected series. This is where the notebook makes its analytical choice explicit: the parameters chosen here determine what kind of pattern the detector or transformer will become sensitive to and how the later outputs should be interpreted.

```{r}
# Configure and fit the ARIMA-based detector
model <- hanr_arima()
model <- fit(model, dataset$serie)
```







### Run the Core Analysis

With the environment and the method ready, we execute the central analytical step and inspect its immediate output. This is the point where the abstract idea described earlier becomes operational, so the reader should pay attention to what is produced and how Harbinger standardizes the result.

```{r}
# Detect anomalies from model residuals
detection <- detect(model, dataset$serie)
head(detection)
```







### Evaluate What Was Found

After producing detections or transformed outputs, we compare them with the reference labels whenever they are available. This stage matters because it connects the visual intuition of the method with an explicit measurement of quality, helping the learner understand not only whether the method runs, but how well it behaves.

```{r}
# Evaluate against the ground truth
evaluation <- evaluate(model, detection$event, dataset$event)
evaluation$confMatrix
```







### Interpret the Result Visually

The final plots are not just illustrations. They help the reader connect the method's internal output with the original series, making it easier to see why a point, range, motif, or symbolic pattern was emphasized and whether that emphasis is coherent with the stated objective of the example.

```{r}
# Plot detections on the original series
har_plot(model, dataset$serie, detection, dataset$event)
```



```{r}
# Plot the residual signal used by the detector
har_plot(
  model,
  attr(detection, "res"),
  detection,
  dataset$event,
  yline = attr(detection, "threshold")
)
```

## References

- Box, G. E. P., Jenkins, G. M., Reinsel, G. C., Ljung, G. M. (2015). Time Series Analysis: Forecasting and Control. Wiley.

