## Objective

This tutorial introduces a model-based anomaly detector using `hanr_arima()`. The goal is to show how Harbinger can detect unusual behavior by modeling expected dynamics and then analyzing residual deviations.

## Method at a glance

`hanr_arima()` fits an ARIMA model to the series, computes residual magnitudes, and flags unusually large residuals as anomalies. This moves from a simple distribution-based baseline to a temporal model that tries to explain normal behavior first.

## What you will do

- fit an ARIMA-based detector
- run anomaly detection on a labeled series
- inspect the evaluation output
- plot residual magnitudes and thresholds

## Walkthrough

```{r}
library(daltoolbox)
library(harbinger)
```

```{r}
data(examples_anomalies)
dataset <- examples_anomalies$simple
```

```{r}
# Configure and fit the ARIMA-based detector
model <- hanr_arima()
model <- fit(model, dataset$serie)
```

```{r}
# Detect anomalies from model residuals
detection <- detect(model, dataset$serie)
head(detection)
```

```{r}
# Evaluate against the ground truth
evaluation <- evaluate(model, detection$event, dataset$event)
evaluation$confMatrix
```

```{r}
# Plot detections on the original series
har_plot(model, dataset$serie, detection, dataset$event)
```

```{r}
# Plot the residual signal used by the detector
har_plot(
  model,
  attr(detection, "res"),
  detection,
  dataset$event,
  yline = attr(detection, "threshold")
)
```

## References

- Box, G. E. P., Jenkins, G. M., Reinsel, G. C., Ljung, G. M. (2015). Time Series Analysis: Forecasting and Control. Wiley.
