## Objective

This tutorial explains how to evaluate detected events in Harbinger. The goal is to compare strict event matching with a softer evaluation that tolerates small timing deviations.

## Method at a glance

Harbinger supports both hard and soft evaluation. Hard evaluation requires the detected event to match the labeled event position directly. Soft evaluation relaxes this and can be more appropriate when a detector finds the right phenomenon but marks it a few timestamps earlier or later.

## What you will do

- run an anomaly detector
- compute hard evaluation metrics
- compute soft evaluation metrics
- compare the two interpretations

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough







### Prepare the Example

We begin by organizing the environment, loading the packages, and selecting the dataset used in the notebook. This part is intentionally more direct: the goal is to make the starting point explicit before the method-specific reasoning begins.

```{r}
library(daltoolbox)
library(harbinger)
```



### Configure the Method

The next step is to instantiate the method and, when necessary, fit it to the selected series. This is where the notebook makes its analytical choice explicit: the parameters chosen here determine what kind of pattern the detector or transformer will become sensitive to and how the later outputs should be interpreted.

```{r}
data(examples_anomalies)
dataset <- examples_anomalies$simple
model <- hanr_arima()
model <- fit(model, dataset$serie)
detection <- detect(model, dataset$serie)
```







### Evaluate What Was Found

After producing detections or transformed outputs, we compare them with the reference labels whenever they are available. This stage matters because it connects the visual intuition of the method with an explicit measurement of quality, helping the learner understand not only whether the method runs, but how well it behaves.

```{r}
# Hard evaluation: direct match between detected and labeled events
hard_eval <- har_eval()
hard_result <- evaluate(hard_eval, detection$event, dataset$event)
hard_result$confMatrix
```



```{r}
# Soft evaluation: tolerates small temporal misalignment
soft_eval <- har_eval_soft()
soft_result <- evaluate(soft_eval, detection$event, dataset$event)
soft_result$confMatrix
```







```{r}
# Inspect summary metrics returned by the evaluation objects
hard_result
soft_result
```

## References

- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3

