## Objective

This tutorial explains how to evaluate detected events in Harbinger. The goal is to compare strict event matching with a softer evaluation that tolerates small timing deviations.

## Method at a glance

Harbinger supports both hard and soft evaluation. Hard evaluation requires the detected event to match the labeled event position directly. Soft evaluation relaxes this and can be more appropriate when a detector finds the right phenomenon but marks it a few timestamps earlier or later.

## What you will do

- run an anomaly detector
- compute hard evaluation metrics
- compute soft evaluation metrics
- compare the two interpretations

## Walkthrough

```{r}
library(daltoolbox)
library(harbinger)
```

```{r}
data(examples_anomalies)
dataset <- examples_anomalies$simple
model <- hanr_arima()
model <- fit(model, dataset$serie)
detection <- detect(model, dataset$serie)
```

```{r}
# Hard evaluation: direct match between detected and labeled events
hard_eval <- har_eval()
hard_result <- evaluate(hard_eval, detection$event, dataset$event)
hard_result$confMatrix
```

```{r}
# Soft evaluation: tolerates small temporal misalignment
soft_eval <- har_eval_soft()
soft_result <- evaluate(soft_eval, detection$event, dataset$event)
soft_result$confMatrix
```

```{r}
# Inspect summary metrics returned by the evaluation objects
hard_result
soft_result
```

## References

- Ogasawara, E., Salles, R., Porto, F., Pacitti, E. Event Detection in Time Series. Springer, 2025. doi:10.1007/978-3-031-75941-3
