## Objective

This tutorial introduces change-point detection with `hcp_amoc()`. The goal is to show how Harbinger handles structural changes, which differ from isolated anomalies because the objective is to locate a transition in the behavior of the series.

## Method at a glance

`hcp_amoc()` searches for a single most significant change point. This is a good first method for learning because it is simple, focused, and works well as an entry point before more complex segmentation techniques.

## What you will do

- load a change-point dataset
- run a single-change detector
- inspect the detected location
- plot the result against the labeled event

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
library(harbinger)
```

```{r}
data(examples_changepoints)
dataset <- examples_changepoints$simple
head(dataset)
```

```{r}
# Plot the original series with the known change point
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

```{r}
# Configure and run the AMOC detector
model <- hcp_amoc()
detection <- detect(model, dataset$serie)
detection
```

```{r}
# Plot detections against the labeled event
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Killick, R., Eckley, I. A. changepoint: An R Package for Changepoint Analysis. Journal of Statistical Software, 58(3), 2014.

