## Tutorial 07 - First Change-Point Detector

Change-point detection is different from point-anomaly detection: the target is a transition in the behavior of the series, not only an isolated unusual observation. `hcp_amoc()` is a good first method because it searches for one dominant break.

This notebook keeps the workflow short so the reader can focus on the meaning of a labeled change point.

The technique behind AMOC is single-break optimization: among all possible split positions, the method searches for the location that best separates two regimes with distinct statistical behavior. It is a strong entry point before studying multiple-break methods.

```{r}
library(harbinger)

data(examples_changepoints)
dataset <- examples_changepoints$simple
head(dataset)
```

Plot the original series and the labeled change point.
```{r}
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

Configure the detector and inspect the detected location.
```{r}
model <- hcp_amoc()
detection <- detect(model, dataset$serie)
detection
```

Plot the result against the reference event.
```{r}
har_plot(model, dataset$serie, detection, dataset$event)
```

## References

- Killick, R., Eckley, I. A. changepoint: An R Package for Changepoint Analysis. Journal of Statistical Software, 58(3), 2014.
