## Tutorial 08 - Smoothing with MAS

Transformations matter because they change what becomes visible in the series. `mas()` computes a moving average, which reduces local variation and can expose broader movement patterns.

This tutorial compares two window sizes so the reader can see the trade-off between preserving detail and producing a smoother signal.

The technique is moving-average smoothing. Each point in the transformed series summarizes a local neighborhood, which suppresses short-term noise and highlights lower-frequency structure that may be easier to inspect or model.

```{r}
library(harbinger)

data(examples_changepoints)
dataset <- examples_changepoints$simple
```

Plot the original signal first.
```{r}
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

Compute moving averages with two window sizes.
```{r}
ma_5 <- mas(dataset$serie, order = 5)
ma_15 <- mas(dataset$serie, order = 15)
```

The smaller window keeps more local variation.
```{r}
har_plot(harbinger(), as.numeric(ma_5), event = dataset$event[5:length(dataset$event)])
```

The larger window produces a smoother series.
```{r}
har_plot(harbinger(), as.numeric(ma_15), event = dataset$event[15:length(dataset$event)])
```

## References

- Shumway, R. H., Stoffer, D. S. Time Series Analysis and Its Applications. Springer.
