## Objective

This tutorial shows how `mas()` can be used as a transformation before a modeling step. The goal is to understand how smoothing removes part of the local variation and exposes broader movement patterns in the series.

## Method at a glance

`mas()` computes a simple moving average. A short window keeps more local detail, while a larger window produces a smoother signal. This is useful before visual inspection, feature extraction, or downstream detection.

## What you will do

- load a time series
- compute moving averages with two window sizes
- compare the original and transformed versions
- interpret the effect of the smoothing window

## How to read this walkthrough

The code blocks below follow the same learning rhythm used throughout the collection: prepare the environment, choose the dataset, configure the method, run the analysis, and then inspect the result. Readers who are still learning time-series mining can use that order to understand not only *what* each command does, but also *why* it appears at that stage of the workflow.

As you go through the notebook, read the inline comments inside each chunk as the operational explanation and use the surrounding prose as the conceptual guide.

## Walkthrough

```{r}
library(harbinger)
```

```{r}
data(examples_changepoints)
dataset <- examples_changepoints$simple
```

```{r}
# Original signal
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

```{r}
# Two smoothing levels
ma_5 <- mas(dataset$serie, order = 5)
ma_15 <- mas(dataset$serie, order = 15)
```

```{r}
# Plot the smoother with a smaller window
har_plot(harbinger(), as.numeric(ma_5), event = dataset$event[5:length(dataset$event)])
```

```{r}
# Plot the smoother with a larger window
har_plot(harbinger(), as.numeric(ma_15), event = dataset$event[15:length(dataset$event)])
```

## References

- Shumway, R. H., Stoffer, D. S. Time Series Analysis and Its Applications. Springer.

