## Objective

This tutorial introduces symbolic transformations for time series with `trans_sax()` and `trans_xsax()`. The goal is to show how a numeric series can be re-encoded into symbols before motif or discord analysis.

## Method at a glance

SAX and XSAX discretize the series into symbolic codes. SAX uses a compact alphabet, while XSAX allows a larger one. Both can simplify comparison and highlight shape patterns that are easier to study symbolically than numerically.

## What you will do

- load a motif example series
- transform the signal with SAX
- transform the same signal with XSAX
- compare the first symbolic values returned by each method

## Walkthrough

```{r}
library(daltoolbox)
library(harbinger)
```

```{r}
data(examples_motifs)
dataset <- examples_motifs$simple
har_plot(harbinger(), dataset$serie, event = dataset$event)
```

```{r}
# SAX uses a compact symbolic alphabet
sax_model <- trans_sax(alpha = 8)
sax_model <- fit(sax_model, dataset$serie)
sax_series <- transform(sax_model, dataset$serie)
head(sax_series, 20)
```

```{r}
# XSAX uses a larger alphabet for finer symbolic resolution
xsax_model <- trans_xsax(alpha = 16)
xsax_model <- fit(xsax_model, dataset$serie)
xsax_series <- transform(xsax_model, dataset$serie)
head(xsax_series, 20)
```

```{r}
# Compare the symbolic encodings side by side
head(
  data.frame(
    value = dataset$serie,
    sax = sax_series,
    xsax = xsax_series
  ),
  20
)
```

## References

- Lin, J., Keogh, E., Lonardi, S., Chiu, B. (2007). A symbolic representation of time series, with implications for streaming algorithms. Data Mining and Knowledge Discovery, 15, 107-144.
